---
title: "Demographic Summary for Bacterial Pneumonia Comparing Cases to Controls"
author: "Dave Bridges"
date: "June 18, 2024"
output:
  html_document:
    highlight: tango
    keep_md: yes
    number_sections: yes
    toc: yes
---

## Purpose

This script combines the cleaned datasets of cases of bactrial pneumonia and controls with no types of pneumonia and compares demographic characteristics.  This script can be found in `r getwd()` and was most recently run on `r date()`.

```{r input-data}
library(knitr)
#figures made will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures-case-control/',
               echo=TRUE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)

library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(lubridate)

control.filename <- 'DatasetComplete.csv'
control.data <- read_csv(control.filename) %>%
  mutate(Class="Controls") %>%
  rename("PriorCardiacArrhythmias"="CardiacArrhythmias",
         "PriorCOPD"="COPD",
         "PriorDiabetes"="Diabetes",
         "PriorHypertension"="Hypertension",
         "PriorObesity"="Obesity")

case.filename <- '../cases/DatasetComplete.csv'
case.data <- read_csv(case.filename) %>%
  filter(Type=="Bacterial") %>%
  arrange(AdmitDate) %>%
  distinct(DeID_PatientID,.keep_all = T) %>%
  mutate(Class="Cases") %>%
  select(-ends_with('First')) %>%
  rename("AgeInYears"="AgeInYears.x") 

combined.data.initial <-
  bind_rows(case.data,control.data)

genotype.datafile <- '../Q223R.txt'
genotype.data <- read_tsv(genotype.datafile)

combined.data %>%
  filter(DeID_PatientID %in% genotype.data$DeID_PatientID)
```

Initially there were `r nrow(combined.data.initial)` patients, but after filtering for those with *LEPR* Q223R genotype data there was `r nrow(combined.data)` patients, including `r nrow(combined.data %>% filter(Class=="Cases"))` cases and `r nrow(combined.data %>% filter(Class=="Controls"))` controls.`

# Case vs Control

```{r case-control}
quant.demo <- 
  combined.data %>%
  group_by(Class) %>%
  summarize(across(.cols=c('AgeInYears','BMI','MaxStay'),
                   .fns=list(mean = ~mean(.,na.rm=T),
                             sd = ~sd(.,na.rm=T),
                             #shapiro.p = ~shapiro.test(.)$p.value,
                             n = ~length(!(is.na(.)))))) %>%
  pivot_longer(cols=contains("_"), #fix this
               names_sep="_",names_to=c("Variable","Statistic")) %>%
  pivot_wider(names_from=Statistic,values_from=value) %>%
  arrange(Variable)

kable(quant.demo, caption="Summary of quantitative values")
# wilcoxon tests, not normally distributed

# quant.t.tests <-
#    combined.data %>%
#   mutate(Class=as.factor(Class)) %>%
#   summarize(across(.cols=c('AgeInYears','BMI'), 
#                    .fns=list(Mann.Whitney=~wilcox.test(.~Class)$p.value)))
# 
# kable(quant.t.tests,caption="Mann-Whitney tests for 30 day survival",digits=c(99,99,99))

ances.demo.ql <- 
  combined.data %>%
  group_by(MajorityAncestry,Class) %>%
  mutate(Type="Ancestry") %>%
  rename("Group"="MajorityAncestry") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ancestry")

type.demo.ql <- 
  combined.data %>%
  group_by(Type,Class) %>%
  rename("Group"="Type") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Pneumonia Type")

emerg.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.x,Class) %>%
  rename("Group"="EmergencyVisit.x") %>%
  count %>%
  mutate(Group=as.factor(Group)) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Emergency Visit") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

# coded as dates not yes/no
# cvd.demo.ql <- 
#   combined.data %>%
#   group_by(CerebrovascularDiseaseFirst) %>%
#   rename("Group"="CerebrovascularDiseaseFirst") %>%
#   count %>%
#   ungroup %>% 
#   mutate(Group=as.factor(Group)) %>%
#   mutate(Pct=n/sum(n)*100) %>%
#   mutate(Type="Cerebrovascular Disease")

gender.demo.ql <- 
  combined.data %>%
  group_by(GenderCode,Class) %>%
  rename("Group"="GenderCode") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Gender")

race.demo.ql <- 
combined.data %>%
  group_by(RaceName,Class) %>%
  count %>%
  ungroup %>%
  rename("Group"="RaceName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Race")

ethnicity.demo.ql <- 
combined.data %>%
  group_by(EthnicityName,Class) %>%
  count %>%
  ungroup %>%
  rename("Group"="EthnicityName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ethnicity")

diabetes.demo.ql <- 
  combined.data %>%
  group_by(PriorDiabetes,Class) %>%
  count %>%
  ungroup %>%
  rename("Group"="PriorDiabetes") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorDiabetes") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

smoking.demo.ql <- 
  combined.data %>%
  group_by(SmokingStatusMapped,Class) %>%
  #mutate(Type="Smoking") %>%
  rename("Group"="SmokingStatusMapped") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Smoking") 

#update to hypertension first
hypertension.demo.ql <- 
  combined.data %>%
  group_by(PriorHypertension,Class) %>%
  rename("Group"="PriorHypertension") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorHypertension") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

arrythmia.demo.ql <- 
  combined.data %>%
  group_by(PriorCardiacArrhythmias,Class) %>%
  rename("Group"="PriorCardiacArrhythmias") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Cardiac Arrhythmias") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

obesity.demo.ql <- 
  combined.data %>%
  group_by(PriorObesity,Class) %>%
  rename("Group"="PriorObesity") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Obesity") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

copd.demo.ql <- 
  combined.data %>%
  group_by(PriorCOPD,Class) %>%
  rename("Group"="PriorCOPD") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior COPD") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

death.demo.ql <- 
  combined.data %>%
  group_by(is.na(DeID_DeathIndexDeceasedDate),Class) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Death")  %>%
  mutate(Group=case_when(Group==TRUE~"Deceased",
                         Group==FALSE~"Alive"))

surv30d.demo.ql <- 
  combined.data %>%
  group_by(Survival.30day,Class) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="30d Survival") 

surv60d.demo.ql <- 
  combined.data %>%
  group_by(Survival.60day,Class) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="60d Survival") 

total.survival <- combined.data %>% group_by(Class) %>% count() %>% pull(n)

summary.discrete <-
  bind_rows(gender.demo.ql,
            race.demo.ql,
            ethnicity.demo.ql,
            ances.demo.ql,
            smoking.demo.ql,
            type.demo.ql,
            #emerg.demo.ql, 
            death.demo.ql,
            surv30d.demo.ql,
            surv60d.demo.ql,
            diabetes.demo.ql,
            hypertension.demo.ql,
            arrythmia.demo.ql,
            copd.demo.ql,
            obesity.demo.ql
            ) %>%
  select(Class,Type,Group,n,Pct) %>%
  ungroup %>%
  select(-Pct) %>%
  pivot_wider(names_from = Class,values_from = n, values_fn = sum) %>%
  arrange(Type,Group) %>%
  replace(.=="NULL", NA) %>%
  mutate(Cases = case_when(is.na(Cases)~0,
                            !(is.na(Cases))~Cases)) %>%
  mutate(Controls = case_when(is.na(Controls)~0,
                            !(is.na(Controls))~Controls)) %>%
  rowwise %>%
  mutate(Pct=Cases/(Cases+Controls)*100) %>%
  mutate(Chisq.p = chisq.test(x=c(Cases,Controls),p=total.survival,rescale.p = T)$p.value) %>%
  mutate(Sig = case_when(Chisq.p < 0.01 ~ '**',
                         Chisq.p < 0.05 ~ '*'))

kable(summary.discrete, 
      caption=paste("Summary of discrete variables for bacterial pneumonia cases, compared to controls.  Overall", round(total.survival[1]/(total.survival[2]+total.survival[1])*100,2),"% cases"),
      digits=c(0,0,0,0,1,99,0))
```

# Diabetes and Obesity Interaction

```{r diab-obesty}
library(forcats)
combined.data %>%
  group_by(PriorDiabetes,PriorObesity,Class) %>%
  count() %>%
  ungroup %>%
  pivot_wider(names_from=Class,values_from=n) %>%
  filter(PriorObesity!=-Inf) %>%
  mutate(Pct=Cases/(Cases+Controls)*100) %>%
  mutate(PriorObesity = fct_recode(as.factor(PriorObesity),
                                   "Yes"="1",
                                   "No"="0")) %>%
    mutate(PriorDiabetes = fct_recode(as.factor(PriorDiabetes),
                                   "Yes"="1",
                                   "No"="0"))-> diab.obesity.summary

kable(diab.obesity.summary, caption="Bacterial pneumonia stratified by diabetes or obesity")

library(ggplot2)

ggplot(diab.obesity.summary,aes(y=Pct,x=as.factor(PriorObesity),fill=as.factor(PriorDiabetes))) +
  geom_bar(stat='identity', position='dodge') +
  labs(y="Proportion of Bacterial Pneumonia",
       x="Pre-Existing Obesity",
       fill="Pre-Existing Diabetes") +
  theme_classic(base_size=16) +
  theme(legend.position=c(0.7,0.92))
```

# Output File for Genetic Analysis

```{r output}
genetics_outfile <- '../Case_Control_Pneumonia_Incidence.csv'

combined.data %>%
  mutate(CaCo_Incidence = if_else(Class=="Cases",1,0)) %>%
  mutate(Age = case_when(is.na(Age)~AgeInYears,.default=Age)) |>
  mutate(AgeOver65 = case_when(Age>=65~1,
                               Age<65~0)) %>%
  mutate(COPD.History = case_when(COPD==1~1,
                                  .default=0)) %>%
  select(DeID_PatientID,starts_with("CaCo"),Age,AgeOver65,GenderCode,BMI,SmokingStatusMapped,COPD.History,Diabetes,RaceName,EthnicityName) %>% 
  write_csv(genetics_outfile)


genetics_outfile <- '../Case_Control_Klebsiella_Pneumonia_Incidence.csv'

combined.data %>%
  mutate(CaCo_Incidence = if_else(Class=="Cases"&SubType =='Pneumococcal',1,0)) %>%
  mutate(Age = case_when(is.na(Age)~AgeInYears,.default=Age)) |>
  mutate(AgeOver65 = case_when(Age>=65~1,
                               Age<65~0)) %>%
  mutate(COPD.History = case_when(COPD==1~1,
                                  .default=0)) %>%
  select(DeID_PatientID,starts_with("CaCo"),Age,AgeOver65,GenderCode,BMI,SmokingStatusMapped,COPD.History,Diabetes,RaceName,EthnicityName) %>% 
  write_csv(genetics_outfile)

genetics_outfile <- '../Case_Control_Pneumococcal_Pneumonia_Incidence.csv'

combined.data %>%
  mutate(CaCo_Incidence = if_else(Class=="Cases"&SubType=="Klebsilla",1,0)) %>%
  mutate(Age = case_when(is.na(Age)~AgeInYears,.default=Age)) |>
  mutate(AgeOver65 = case_when(Age>=65~1,
                               Age<65~0)) %>%
  mutate(COPD.History = case_when(COPD==1~1,
                                  .default=0)) %>%
  select(DeID_PatientID,starts_with("CaCo"),Age,AgeOver65,GenderCode,BMI,SmokingStatusMapped,COPD.History,Diabetes,RaceName,EthnicityName) %>% 
  write_csv(genetics_outfile)
```

# Session Information

```{r sessioninfo}
sessionInfo()
```