---
title: "Propensity Matching for Pneumonia Incidence"
author: "Dave Bridges"
date: "2025-12-01"
editor: source
format: 
  html:
    toc: true
    toc-location: right
    keep-md: true
    code-fold: true
    code-summary: "Show the code"
    fig-path: "figures/"
  gfm:
    html-math-method: webtex
theme: journal
execute:
  echo: true
  warning: false
---


```{r global_options}
# hide this code chunk
#| echo: false
#| message: false

# defines the se function
se <- function(x) {
  sd(x, na.rm = TRUE) / sqrt(length(x))
}

#load these packages, nearly always needed
library(tidyverse)
library(knitr)

# sets maize and blue color scheme
color_scheme <- c("#00274c", "#ffcb05")
```

## Data Entry

```{r data-entry}
library(readr)
filename <- 'Case_Control_Pneumonia_Incidence.csv'
data <- read_csv(filename) |>
  mutate(Race.Ethnicity = case_when(RaceName=="African American" ~ "Black",
                                    RaceName=="Caucasian"~"White",
                                    RaceName=="Asian"~"Asian",
                                    EthnicityName=="Hispanic or Latino"~"Hispanic or Latino",
                                    is.na(RaceName)~"Other",
                                    is.na(EthnicityName)~"Other",
                                    .default="Other")) 

library(janitor)

tabyl(data, Race.Ethnicity,CaCo_Incidence)|>
  adorn_percentages("col") %>%          # row percentages (each row sums to 100%)
  adorn_pct_formatting(digits = 0) %>%  # show as 23% instead of 0.23
  adorn_ns(position = "front") %>%      # puts counts in front: 45 (23%)
  adorn_title("combined") |> kable(caption="Counts by race/ethnicity prior to matching")

tabyl(data, GenderCode,CaCo_Incidence)|>
  adorn_percentages("col") %>%          # row percentages (each row sums to 100%)
  adorn_pct_formatting(digits = 0) %>%  # show as 23% instead of 0.23
  adorn_ns(position = "front") %>%      # puts counts in front: 45 (23%)
  adorn_title("combined") |> kable(caption="Counts by gender prior to matching")

data |>
  group_by(CaCo_Incidence) |>
  summarise(Mean = mean(Age,na.rm=T),
            SD = sd(Age, na.rm=T),
            n = length(Age))|> 
  kable(caption="mean age by gender prior to matching")
```

Loaded in case-control unmatched file `r filename`.  This input file was most recently updated on `r ifelse(filename %in% dir(), format(file.info(filename)$mtime, "%Y-%m-%d"), "unknown")`.  This script was most recently updated on `r date()`.

## Propensity Matching

The goal is to try to match cases by age, race, ethnicity and gender, with exact matches on the last three.

```{r matching}
library(MatchIt)
m.out.all <- matchit(CaCo_Incidence ~ Age, 
                 data = data |> filter(is.finite(Age), !is.na(Race.Ethnicity)),
                 method = "nearest", 
                 exact = c("GenderCode", "Race.Ethnicity"),# Force exact matches
                 ratio=3)

matched_data <- match.data(m.out.all)

#summary(m.out.all, standardize = TRUE)  # Balance statistics[5]

tabyl(matched_data, Race.Ethnicity,CaCo_Incidence)|>
  adorn_percentages("col") %>%          # row percentages (each row sums to 100%)
  adorn_pct_formatting(digits = 0) %>%  # show as 23% instead of 0.23
  adorn_ns(position = "front") %>%      # puts counts in front: 45 (23%)
  adorn_title("combined") |> kable(caption="Counts by race/ethnicity after matching")

tabyl(matched_data, GenderCode,CaCo_Incidence)|>
  adorn_percentages("col") %>%          # row percentages (each row sums to 100%)
  adorn_pct_formatting(digits = 0) %>%  # show as 23% instead of 0.23
  adorn_ns(position = "front") %>%      # puts counts in front: 45 (23%)
  adorn_title("combined") |> kable(caption="Counts by gender after matching")

matched_data |>
  group_by(CaCo_Incidence) |>
  summarise(Mean = mean(Age,na.rm=T),
            SD = sd(Age, na.rm=T),
            n = length(Age))|> 
  kable(caption="mean age by gender after matching")
```

### Love Plot

```{r love-plots}
library(cobalt)
library(ggplot2)

#plot(m.out.all,type="jitter")

love.plot(m.out.all, stats="mean.diffs",
          threshold=0.1,
          abs=TRUE,
          stars="none") + 
  theme_classic(base_size=16) +
    labs(
      title = "Love Plot",
      x = "Standardized Mean Difference",
      color = "Stage",
    ) +
    scale_color_grey() +
    theme(legend.position = c(0.65,0.25))

write_csv(matched_data, 'Case_Control_Pneumonia_Incidence_Matched.csv')
```

## Session Information

```{r session-information}
sessionInfo()
```
