\documentclass[a4paper,12pt]{letter}

% Some of the article customisations are relevant for this class
\usepackage{times}
\name{Dave Bridges} % To be used for the return address on the envelope
\signature{Dave Bridges\\Assistant Professor\\UTHSC} % Goes after the closing (ie at the end of the letter, with space for a signature)
%\address{Address \\ of \\ Sender}
% Alternatively, these may be set on an individual basis within each letter environment.


\begin{document}
\begin{letter}{Dr. Hari Hundal\\Associate Editor\\AJP Endocrinology and Metabolism}

\opening{Dr. Hundal,} % eg Hello.

Please find enclosed our revised manuscript \textbf{"Exposure to environmentally persistent free radicals during gestation lowers energy expenditure and impairs skeletal muscle mitochondrial function in adult mice"} by Stephenson \textit{et al.} for further consideration by AJP Endocrinology and Metabolism.  We would like this study to be considered under the \emph{Mitochondrial Dynamics and Oxidative Stress in Disease} call for papers.

In this work, we described an important and novel role of mitochondrial dysfunction early in development which resulted in chronic metabolic abnormalities in offspring.  We have considered the reviewers comments and responded to them in the attached document.  Notably we have included substantial new data regarding insulin sensitivity, oxidative stress and the molecular changes associated with our observed changes in energy expenditure.  We are grateful to the efforts of you and the reviewers in helping us improve this manuscript.

Thank you for your consideration and please contact us if we can assist you further in the review process.

\closing{Regards,} % eg Regards,


\end{letter}
\end{document}
