---
title: "Analysis of Body Composition for High Fat Diet Particulate Treatment Study"
author: Alyse Ragauskas, Matt Peloquin, Jyothi Parvathareddy, Sridhar Jaligama, Stephania
  Cormier and Dave Bridges
date: "November 13, 2014"
output:
  html_document:
    keep_md: true
---

This only looks at animals treated *in utero*.  These data were most recently updated on `r date()`.

```{r data-entry}

output_file_maternal <- "../data/Raw Weights - Maternal.csv"

data.maternal <- read.csv(output_file_maternal, row.names="X")
#removed sickly, weight losing mouse 206
data.maternal <- subset(data.maternal, animal.MouseID !=206)
data.maternal$animal.MouseID <- as.factor(data.maternal$animal.MouseID)

#find which experiments we did fat masss
fat.mass.experiments <- droplevels(subset(data.maternal, assay.assay=="Total Fat Mass"))$experiment.date

data <- subset(data.maternal, experiment.date %in% fat.mass.experiments)
data$animal.Cage <- as.factor(data$animal.Cage)
data$age.group <- cut(data$age, breaks=2)
data$Treatment <- relevel(data$Treatment, ref='Saline')
library(reshape2)
composition.data <- dcast(data, 
                          animal.MouseID+Treatment+animal.Cage+age~assay.assay, value.var="values", 
                          fun.aggregate = function(x) mean(x)/1000)
#remove whitespace
colnames(composition.data) <- gsub(" ", ".", colnames(composition.data))
composition.data$Fat.Pct <- composition.data$Total.Fat.Mass/composition.data$Body.Weight*100
composition.data$Lean.Pct <- composition.data$Lean.Mass/composition.data$Body.Weight*100
```


## Animals Reared from Particulate Treated Mothers

```{r summary-statistics}
library(dplyr)
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
                             
fat.mass.summary <- subset(composition.data, age>145) %>%
  group_by(Treatment) %>%
  distinct(animal.MouseID) %>%
  summarise(mean = mean(Total.Fat.Mass),
            se = se(Total.Fat.Mass ),
            sd = sd(Total.Fat.Mass ),
            rel.sd = sd(Total.Fat.Mass)/mean(Total.Fat.Mass)*100,
            n = length(Total.Fat.Mass),
            shapiro = shapiro.test(Total.Fat.Mass)$p.value)

fat.pct.summary <- subset(composition.data, age>145) %>%
  group_by(Treatment) %>%
  distinct(animal.MouseID) %>%
  summarise(mean = mean(Fat.Pct),
            se = se(Fat.Pct),
            sd = sd(Fat.Pct),
            rel.sd = sd(Fat.Pct)/mean(Fat.Pct)*100,
            n = length(Fat.Pct),
            shapiro = shapiro.test(Fat.Pct)$p.value)

lean.mass.summary <- subset(composition.data, age>145) %>%
  group_by(Treatment) %>%
  distinct(animal.MouseID) %>%
  summarise(mean = mean(Lean.Mass),
            se = se(Lean.Mass),
            sd = sd(Lean.Mass),
            rel.sd = sd(Lean.Mass)/mean(Lean.Mass)*100,
            n = length(Lean.Mass),
            shapiro = shapiro.test(Lean.Mass)$p.value)

lean.pct.summary <- subset(composition.data, age>145) %>%
  group_by(Treatment) %>%
  distinct(animal.MouseID) %>%
  summarise(mean = mean(Lean.Pct),
            se = se(Lean.Pct),
            sd = sd(Lean.Pct),
            rel.sd = sd(Lean.Pct)/mean(Lean.Pct)*100,
            n = length(Lean.Pct),
            shapiro = shapiro.test(Lean.Pct)$p.value)
library(car)
```

# After 12 Weeks of High Fat Diet

```{r summary-tables-pups, results='asis'}
library(xtable)
print(xtable(as.data.frame(fat.pct.summary), label="tab:summary-fat-pct", caption="Data for Percent Fat", digits=3), type='html')
print(xtable(as.data.frame(fat.mass.summary), label="tab:summary-fat-mass", caption="Data for Total Fat", digits=3), type='html')
print(xtable(as.data.frame(lean.mass.summary), label="tab:summary-lean-mass", caption="Data for Total Lean Mass", digits=3), type='html')
```

## Total Fat Mass

```{r fat-mass-barplot, dev=c('png','pdf')}
ymax = max(fat.mass.summary$mean + fat.mass.summary$se)
plot <- with(fat.mass.summary, barplot(mean,
                   las=1,
                   ylab ="Total Fat Mass (g)",
                   names.arg=Treatment,
                   ylim = c(0,ymax)))

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

superpose.eb(plot, fat.mass.summary$mean, fat.mass.summary$se)

unique.composition.data <- distinct(subset(composition.data, age>145), animal.MouseID)
```

The data were normally distributed (p>`r min(fat.mass.summary$shapiro)`) and had equal variance via a Levene's test (p=`r leveneTest(Total.Fat.Mass~Treatment, data=unique.composition.data)$"Pr(>F)"[1]`).  Therefore via a Student's *t*-test, the p-value was `r t.test(Total.Fat.Mass~Treatment, data=unique.composition.data, var.equal=T)$p.value`.  There was a `r (fat.mass.summary$mean[2]-fat.mass.summary$mean[1])/fat.mass.summary$mean[1]*100`% increase in fat mass.

# Total Lean Mass

```{r lean-mass-barplot, dev=c('png','pdf')}
ymax = max(lean.mass.summary$mean + lean.mass.summary$se)
plot <- with(lean.mass.summary, barplot(mean,
                   las=1,
                   ylab ="Total Lean Mass (g)",
                   names.arg=Treatment,
                   ylim = c(0,ymax)))
superpose.eb(plot, lean.mass.summary$mean, lean.mass.summary$se)
```

The data were normally distributed (p>`r min(lean.mass.summary$shapiro)`) and had equal variance via a Levene's test (p=`r leveneTest(Lean.Mass~Treatment, data=unique.composition.data)$"Pr(>F)"[1]`).  Therefore via a Student's *t*-test, the p-value was `r t.test(Lean.Mass~Treatment, data=unique.composition.data, var.equal=T)$p.value`.  There was a `r (lean.mass.summary$mean[2]-lean.mass.summary$mean[1])/lean.mass.summary$mean[1]*100`% increase in lean mass.

# Percent Fat Mass
```{r fat-pct-barplot, dev=c('png','pdf')}
ymax = max(fat.pct.summary$mean + fat.pct.summary$se)
plot <- with(fat.pct.summary, barplot(mean,
                   las=1,
                   ylab ="Percent Body Fat",
                   names.arg=Treatment,
                   ylim = c(0,ymax)))

superpose.eb(plot, fat.pct.summary$mean, fat.pct.summary$se)
```


The data were normally distributed (p>`r min(fat.pct.summary$shapiro)`) and had **unequal variance** via a Levene's test (p=`r leveneTest(Fat.Pct~Treatment, data=unique.composition.data)$"Pr(>F)"[1]`).  Therefore via a Welch's *t*-test, the p-value was `r t.test(Fat.Pct~Treatment, data=unique.composition.data, var.equal=F)$p.value`.

## Session Information
```{r sessionInfo}
sessionInfo()
```
