---
title: "Fasting Glucose Analysis for Particulate Treatment Study"
author: Alyse Ragauskas, Matt Peloquin, Jyothi Parvathareddy, Sridhar Jaligama, Stephania
  Cormier and Dave Bridges
date: "November 13, 2014"
output:
  html_document:
    keep_md: true
---

This only looks at animals treated *in utero*.  These data were most recently updated on `r date()`.

```{r data-entry}
filename <- '../data/Blood Glucose - Maternal.csv'
data <- read.csv(filename, row.names='X')
data <- subset(data, animal.MouseID!='206')

library(plyr)
summary <- ddply(data, .(Treatment, experiment.feeding_state), summarize,
                 mean = mean(Glucose, na.rm=T),
                 sd = sd(Glucose, na.rm=T),
                 se = sd(Glucose, na.rm=T)/sqrt(length(Glucose)),
                 n = length(Glucose),
                 shapiro = shapiro.test(Glucose)$p.value)

library(car)
levene.result <- leveneTest(Glucose~Treatment, data=subset(data, experiment.feeding_state=='fasted'))
```

The data is located in the file`r filename`.  These data are summarized in the Table below and graphed in the figure.  


```{r summary-tables, results='asis'}
library(xtable)
print(xtable(summary, caption = "Summary Data, based on treating mice individually.  Triglycerides are in mmoles/mg of tissue", label="tab:summary-statistics"), include.rownames=F, type='html')
```

According to a Shapiro-Wilk Test, the data fit a normal distribution (p>`r min(summary$shapiro)`).  A Levene's test suggested that the variance can be presumed to be equal (p=`r levene.result$"Pr(>F)"[1]`).  Based on this, a Student's T-test has a p-value of `r t.test(Glucose~Treatment, data=subset(data, experiment.feeding_state == 'fasted'), var.equal=T)$p.value`.

```{r barplot-individual,dev=c('png','pdf')}
ymax <- max(summary$mean) + max(summary$se)

plot <- with(subset(summary, experiment.feeding_state == 'fasted'), plot <- barplot(mean,
                   beside=T,
                   las=1,
                   ylab ="Blood Glucose (mg/dL)",
                   names.arg = Treatment,
                   ylim = c(0,ymax)))

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

superpose.eb(plot, subset(summary, 
            experiment.feeding_state == 'fasted')$mean, 
             subset(summary, experiment.feeding_state == 'fasted')$se)
```

## Session Information
```{r sessionInfo}
sessionInfo()
```
