---
title: "Food Intake Analysis for Particulate Treatment Study"
author: Alyse Ragauskas, Matt Peloquin, Jyothi Parvathareddy, Sridhar Jaligama, Stephania
  Cormier and Dave Bridges
date: "November 13, 2014"
output:
  html_document:
    keep_md: true
---

This part of the study involved just one arm, animals treated *in utero*.  This file was most recently updated on `r date()`.

```{r data-entry, echo=FALSE}
input_file <- '../data/Food Worksheet.xlsx'
sheet_used <- 'Maternal-Summary'
library(xlsx)
raw.data <- read.xlsx2(input_file, sheetName=sheet_used, colClasses=c('character','character','numeric','numeric'))
raw.data$Cage <- as.factor(raw.data$Cage)
raw.data$Treatment <- as.factor(raw.data$Treatment)
raw.data$Week <- as.integer(raw.data$Week)

library(plyr)

weekly.summary <- ddply(raw.data, .(Treatment, Week), summarize,
                        mean = mean(Calories),
                        sd = sd(Calories),
                        se = sd(Calories)/sqrt(length(Calories)),
                        n = length(Calories))

cumulative.data <- ddply(raw.data, .(Cage,Treatment), transform,
                                   Cumulative.Calories = cumsum(Calories))

weekly.summary.cumulative  <- ddply(cumulative.data, .(Treatment, Week), summarize,
                        mean = mean(Cumulative.Calories),
                        sd = sd(Cumulative.Calories),
                        se = sd(Cumulative.Calories)/sqrt(length(Cumulative.Calories)),
                        n = length(Cumulative.Calories))
```

This data comes from the file `r input_file` using the worksheet `r sheet_used`.  

## Animals Treated with Particulate Matter Gestationally

```{r weekly-food-intake,fig=TRUE, dev=c('png','pdf')}
ymax <- max(weekly.summary$se + weekly.summary$mean)
plot <- with(subset(weekly.summary, Treatment=="Control"), plot(Week, mean,
             pch=18, las=1, ylab="Calories per mouse per week", ylim=c(0,ymax)))

with(subset(weekly.summary, Treatment=="Particulate"), points(Week, mean, pch=18, col="red"))


superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

#lines
abline(lm(mean~Week, data=subset(weekly.summary, Treatment=="Control")), col="black")
abline(lm(mean~Week, data=subset(weekly.summary, Treatment=="Particulate")), col="red")

#error bars
superpose.eb(subset(weekly.summary, Treatment=="Control")$Week,
             subset(weekly.summary, Treatment=="Control")$mean,
             subset(weekly.summary, Treatment=="Control")$se)

superpose.eb(subset(weekly.summary, Treatment=="Particulate")$Week,
             subset(weekly.summary, Treatment=="Particulate")$mean,
             subset(weekly.summary, Treatment=="Particulate")$se, col="red")

legend("bottomleft", levels(raw.data$Treatment), col=c("black",'red'), pch=18, bty="n", lty=1)
```

```{r weekly-food-intake-cumulative,echo=FALSE, dev=c('png','pdf')}
ymax <- max(weekly.summary.cumulative$se + weekly.summary.cumulative$mean)
plot <- with(subset(weekly.summary.cumulative, Treatment=="Control"), plot(Week, mean,
             pch=18, las=1, ylab="Cumulative Calories Per Mouse", ylim=c(0,ymax)))

with(subset(weekly.summary.cumulative, Treatment=="Particulate"), points(Week, mean, pch=18, col="red"))

#lines
abline(lm(mean~Week-1, data=subset(weekly.summary.cumulative, Treatment=="Control")), col="black")
abline(lm(mean~Week-1, data=subset(weekly.summary.cumulative, Treatment=="Particulate")), col="red")

#error bars
superpose.eb(subset(weekly.summary.cumulative, Treatment=="Control")$Week,
             subset(weekly.summary.cumulative, Treatment=="Control")$mean,
             subset(weekly.summary.cumulative, Treatment=="Control")$se)

superpose.eb(subset(weekly.summary.cumulative, Treatment=="Particulate")$Week,
             subset(weekly.summary.cumulative, Treatment=="Particulate")$mean,
             subset(weekly.summary.cumulative, Treatment=="Particulate")$se, col="red")

legend("topleft", levels(raw.data$Treatment), col=c("black",'red'), pch=18, bty="n", lty=1)
```


```{r statistics, echo=FALSE}
library(lme4)
food.lme <- lmer(Calories ~ Treatment*Week + (1|Cage), data=raw.data, REML=T)
food.lme.no.treatment <- lmer(Calories ~ Week + (1|Cage), data=raw.data, REML=T)

food.lme.cumulative <- lmer(Cumulative.Calories ~ Week + Treatment:Week -1 + (1|Cage), data=cumulative.data, REML=T)
food.lme.cumulative.no.treatment <- lmer(Cumulative.Calories ~ Week -1 + (1|Cage), data=cumulative.data, REML=T)
```

We tested by &Chi;2-test whether there is a difference between mixed models considering the amount of food eaten and the treatment.  The residuals of this model fit a normal distribution (p=`r shapiro.test(residuals(food.lme))$p.value` from a Shapiro-Wilk Test).  The p-value for that test is &Chi;2 =`r anova(food.lme, food.lme.no.treatment)$Chisq[2]`, p=`r anova(food.lme, food.lme.no.treatment)$"Pr(>Chisq)"[2]` so there is no evidence of a difference between the treatments.

If we look at the cumulative food intake, the residuals are still normally distributed (p=`r shapiro.test(residuals(food.lme.cumulative))$p.value` from Shapiro-Wilk Test).  Comparing mixed linear models with an interaction term, and the slope forced to zero there was a significant difference between the two groups by &Chi;2 = `r anova(food.lme.cumulative, food.lme.cumulative.no.treatment)$Chisq[2]` and **p=`r anova(food.lme.cumulative, food.lme.cumulative.no.treatment)$"Pr(>Chisq)"[2]`**.  Based on this, the control group ate `r fixef(food.lme.cumulative)['Week:TreatmentControl']` kcal of food more per mouse per week, or a total of `r fixef(food.lme.cumulative)['Week:TreatmentControl']*12`, kcal throughout the diet.  This corresponds to a decrease of `r fixef(food.lme.cumulative)['Week:TreatmentControl']/(fixef(food.lme.cumulative)['Week:TreatmentControl']+fixef(food.lme.cumulative)['Week'])*100`% of caloric intake

## Session Information
```{r sessionInfo,echo=F}
sessionInfo()
```
