---
title: "Analysis of Fasted Liver Triglycerides from High Fat Diet Particulate Treatment Study"
author: Kathryn Cyrus, Matt Peloquin, Jyothi Parvathareddy, Sridhar Jaligama, Stephania
  Cormier and Dave Bridges
date: "November 13, 2014"
output:
  html_document:
    keep_md: true
---

This only looks at animals treated *in utero* and these values are from mice which were fasted 16-20h.  These data were most recently updated on `r date()`.

```{r data-entry}
filename <- '../data/Liver Triglycerides.xlsx'
worksheet <- 'Summary'

library(xlsx)
data <- read.xlsx2(filename, sheetName=worksheet)
data$TG <- as.numeric(as.character(data$TGs..microgram.milligrams.of.tissue.))

library(plyr)
summary <- ddply(data, ~Treatment, summarize,
                 mean = mean(TG),
                 sd = sd(TG),
                 se = sd(TG)/sqrt(length(TG)),
                 n = length(TG),
                 shapiro = shapiro.test(TG)$p.value)

library(car)
levene.result <- leveneTest(TG~Treatment, data=data)
```

The data is located in the file `r filename` on a worksheet named `r worksheet`.  These data are summarized in the Table below and graphed in the Figure.  


```{r summary-tables, results='asis'}
library(xtable)
print(xtable(summary, caption = "Summary Data, based on treating mice individually.  Triglycerides are in mmoles/mg of tissue", label="tab:summary-statistics"), include.rownames=F, type='html')
```

According to a Shapiro-Wilk Test, the data fit a normal distribution (p>`r min(summary$shapiro)`).  A Levene's test suggested that the variance can be presumed to be equal (p=`r levene.result$"Pr(>F)"[1]`).  Based on this, a Student's T-test has a p-value of `r t.test(TG~Treatment, data=data, var.equal=T)$p.value.

```{r barplot-individual,dev=c('png','pdf')}
ymax <- max(summary$mean) + max(summary$se)

plot <- barplot(summary$mean,
                   beside=T,
                   las=1,
                   ylab ="Liver Triglycerides (ug/mg tissue)",
                   names.arg = summary$Treatment,
                   ylim = c(0,ymax))

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

superpose.eb(plot, summary$mean, summary$se)
```

## Session Information
```{r sessionInfo}
sessionInfo()
```


\end{document}
