---
title: "Fasting Luminex Data for Particulate Treatment Study"
author: Erin Stephenson, Alyse Ragauskas, Matt Peloquin, Jyothi Parvathareddy, Sridhar Jaligama, Stephania
  Cormier and Dave Bridges
date: "November 13, 2014"
output:
  html_document:
    keep_md: true
---

# Summary

Pregnant mice were exposed to respiratory particulates or saline. The offspring of these mice were placed on a HFD at 10 wk of age for 12 wk. Blood was collected in heparinized tubes via retro orbital bleed both before and after a 16 hr fast. Samples were allowed to clot over ice before being separated via centrifugation and serum was collected. The values presented here are the results of serum analyses using the BioRad Bio-Plex Pro Mouse Diabetes Panel (8-plex, Cat. No.171-F7001M).This Assay was performed by Erin Stephenson, with the assistance of Alyse Ragauskas. Jordy Saravia calculated the concentrations of each hormone analyzed.     

```{r data-entry}
input_file <- "../data/Luminex Data.xlsx"
sheet_name <- "Sheet2"
library(xlsx)
data <- read.xlsx2(input_file, sheetName=sheet_name)
#set Saline to be the reference treatment
data$Treatment <- relevel(data$Treatment, ref='Saline')
#remove sick mouse #206
data <- subset(data, Mouse !='206')
data$Resistin <- as.numeric(as.character(data$'Resistin'))
data$GIP <- as.numeric(as.character(data$'GIP'))
data$PAI1 <- as.numeric(as.character(data$'PAI1'))
data$GLP1 <- as.numeric(as.character(data$'GLP1'))
data$Glucagon <- as.numeric(as.character(data$'Glucagon'))
data$Ghrelin <- as.numeric(as.character(data$'Ghrelin'))
data$Leptin <- as.numeric(as.character(data$'Leptin'))
data$Insulin <- as.numeric(as.character(data$'Insulin'))
data$Insulin.pmol.L <- as.numeric(as.character(data$'Insulin.pmol.L'))
#set color palette
Treatment.colors <- rep(c('Black','Red'),2)
```


# Data Summary
```{r summary}
library(plyr)
#define se
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))

Resistin.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(Resistin, na.rm=T),
                     se = se(Resistin),
                     sd = sd(Resistin, na.rm=T),
                     rel.sd = sd(Resistin, na.rm=T)/mean(Resistin, na.rm=T)*100,
                     n = length(Resistin))


GIP.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(GIP, na.rm=T),
                     se = se(GIP),
                     sd = sd(GIP, na.rm=T),
                     rel.sd = sd(GIP, na.rm=T)/mean(GIP, na.rm=T)*100,
                     n = length(GIP))

PAI1.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(PAI1, na.rm=T),
                     se = se(PAI1),
                     sd = sd(PAI1, na.rm=T),
                     rel.sd = sd(PAI1, na.rm=T)/mean(PAI1, na.rm=T)*100,
                     n = length(PAI1))

GLP1.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(GLP1, na.rm=T),
                     se = se(GLP1),
                     sd = sd(GLP1, na.rm=T),
                     rel.sd = sd(GLP1, na.rm=T)/mean(GLP1, na.rm=T)*100,
                     n = length(GLP1))

Glucagon.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(Glucagon, na.rm=T),
                     se = se(Glucagon),
                     sd = sd(Glucagon, na.rm=T),
                     rel.sd = sd(Glucagon, na.rm=T)/mean(Glucagon, na.rm=T)*100,
                     n = length(Glucagon))

Ghrelin.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(Ghrelin, na.rm=T),
                     se = se(Ghrelin),
                     sd = sd(Ghrelin, na.rm=T),
                     rel.sd = sd(Ghrelin, na.rm=T)/mean(Ghrelin, na.rm=T)*100,
                     n = length(Ghrelin))

Leptin.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(Leptin, na.rm=T),
                     se = se(Leptin),
                     sd = sd(Leptin, na.rm=T),
                     rel.sd = sd(Leptin, na.rm=T)/mean(Leptin, na.rm=T)*100,
                     n = length(Leptin))

Insulin.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(Insulin, na.rm=T),
                     se = se(Insulin),
                     sd = sd(Insulin, na.rm=T),
                     rel.sd = sd(Insulin, na.rm=T)/mean(Insulin, na.rm=T)*100,
                     n = length(Insulin))
Insulin.pmol.L.summary <- ddply(data, .(FeedingState,Treatment), summarize,
                     mean = mean(Insulin.pmol.L, na.rm=T),
                     se = se(Insulin.pmol.L),
                     sd = sd(Insulin.pmol.L, na.rm=T),
                     rel.sd = sd(Insulin.pmol.L, na.rm=T)/mean(Insulin.pmol.L, na.rm=T)*100,
                     n = length(Insulin.pmol.L))
```

# Statistics

We first did two way ANOVA's for each of the hormones, with Feeding State and Treatment Group as the main effects and allowing for interactions.  

```{r stats-summary, results='asis'}
library(xtable)

Resistin.aov <- aov(Resistin~FeedingState*Treatment, data=data)
GIP.aov <- aov(GIP~FeedingState*Treatment, data=data)
PAI1.aov <- aov(PAI1~FeedingState*Treatment, data=data)
GLP1.aov <- aov(GLP1~FeedingState*Treatment, data=data)
Glucagon.aov <- aov(Glucagon~FeedingState*Treatment, data=data)
Ghrelin.aov <- aov(Ghrelin~FeedingState*Treatment, data=data)
Leptin.aov <- aov(Leptin~FeedingState*Treatment, data=data)
Insulin.aov <- aov(Insulin~FeedingState*Treatment, data=data)


library(car)
Resistin.levene <- leveneTest(Resistin~FeedingState*Treatment, data=data)
GIP.levene <- leveneTest(GIP~FeedingState*Treatment, data=data)
PAI1.levene <- leveneTest(PAI1~FeedingState*Treatment, data=data)
GLP1.levene <- leveneTest(GLP1~FeedingState*Treatment, data=data)
Glucagon.levene <- leveneTest(Glucagon~FeedingState*Treatment, data=data)
Ghrelin.levene <- leveneTest(Ghrelin~FeedingState*Treatment, data=data)
Leptin.levene <- leveneTest(Leptin~FeedingState*Treatment, data=data)
Insulin.levene <- leveneTest(Insulin~FeedingState*Treatment, data=data)


anova.summary <- data.frame(row.names=colnames(data)[4:11])
for (hormone in rownames(anova.summary)){
  current.anova <- eval(as.symbol(paste(hormone,'aov',sep='.')))
  current.levene <- eval(as.symbol(paste(hormone,'levene',sep='.')))
anova.summary[hormone,'Shapiro'] <- shapiro.test(residuals(current.anova))$p.value
anova.summary[hormone,'Levene'] <- current.levene$`Pr(>F)`[1]
anova.summary[hormone,'FeedingState'] <- summary(current.anova)[[1]]$`Pr(>F)`[1]
anova.summary[hormone,'Treatment'] <- summary(current.anova)[[1]]$`Pr(>F)`[2]
anova.summary[hormone,'Interaction'] <- summary(current.anova)[[1]]$`Pr(>F)`[3]
}

anova.summary.verified <- anova.summary[anova.summary$Shapiro>0.05&anova.summary$Levene>0.05, ]

print(xtable(anova.summary.verified, caption = "Summaries for Two-Way ANOVA Analyses, for the genes under which the ANOVA assumptions were met", digits = 5, label="tab:anova-summary"), type='html')
```

Since the treatment term for the Leptin ANOVA was significant, and both the normality and equal variance assumptions were met, we performed pairwise t-tests for the effects of MCP230 treatment on leptin.  These results are `r  t.test(Leptin~Treatment, data=subset(data,FeedingState=="Fed"), var.equal=T)$p.value` for the fed state and `r  t.test(Leptin~Treatment, data=subset(data,FeedingState=="Fasted"), var.equal=T)$p.value` for the fasted state.


The assumptions of normality were not met for `r paste(rownames( anova.summary[anova.summary$Shapiro<0.05,]), collapse=", ")` and the assumptions of equal variance were not met for `r paste(rownames( anova.summary[anova.summary$Levene<0.05,]), collapse=", ")`.  We therefore did Wilcoxon Rank Sum tests comparing treatment groups, after separating into Fasted and Refed groups.  These are shown in the Table below:

```{r wilcoxon-summary, results='asis'}

wilcoxon.summary <- data.frame(row.names=rownames( anova.summary[anova.summary$Shapiro<0.05,]))

for (hormone in rownames(wilcoxon.summary)){
  wilcoxon.summary[hormone,'Fasted'] <- wilcox.test(formula( paste( hormone, '~ Treatment'))  , data=subset(data, FeedingState=='Fasted'))$p.value
  wilcoxon.summary[hormone,'Fed'] <- wilcox.test(formula( paste( hormone, '~ Treatment')) , data=subset(data, FeedingState=='Fed'))$p.value
}

print(xtable(wilcoxon.summary, caption = "Summaries of Wilcoxon Rank Sum tests, comparing treatment groups separately in fasted and refed states", digits = 5, label="tab:wilcoxon-summary"), type='html')
```

```{r barplot-Resistin, dev=c('png','pdf')}
require(reshape2)
Resistin.summary.means <- as.matrix(dcast(Resistin.summary, Treatment~FeedingState, value.var="mean")[2:3])

Resistin.summary.se <- as.matrix(dcast(Resistin.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Resistin.summary$mean + Resistin.summary$se, na.rm=T)
superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

plot <- barplot(Resistin.summary.means,
                ylab="Serum Resistin Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:4],
                ylim = c(0,ymax),
                beside=T)
legend("topright", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, Resistin.summary.means, Resistin.summary.se)
```

```{r barplot-GIP, dev=c('png','pdf')}
GIP.summary.means <- as.matrix(dcast(GIP.summary, Treatment~FeedingState, value.var="mean")[2:3])

GIP.summary.se <- as.matrix(dcast(GIP.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(GIP.summary$mean + GIP.summary$se, na.rm=T)

plot <- barplot(GIP.summary.means,
                ylab="Serum GIP Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
legend("topleft", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, GIP.summary.means, GIP.summary.se)
```

```{r barplot-PAI1, dev=c('png','pdf')}
PAI1.summary.means <- as.matrix(dcast(PAI1.summary, Treatment~FeedingState, value.var="mean")[2:3])

PAI1.summary.se <- as.matrix(dcast(PAI1.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(PAI1.summary$mean + PAI1.summary$se, na.rm=T)

plot <- barplot(PAI1.summary.means,
                ylab="Serum PAI1 Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
legend("topright", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, PAI1.summary.means, PAI1.summary.se)
```

```{r barplot-GLP1,, dev=c('png','pdf')}
GLP1.summary.means <- as.matrix(dcast(GLP1.summary, Treatment~FeedingState, value.var="mean")[2:3])

GLP1.summary.se <- as.matrix(dcast(GLP1.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(GLP1.summary$mean + GLP1.summary$se, na.rm=T)

plot <- barplot(GLP1.summary.means,
                ylab="Serum GLP1 Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
legend("topleft", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, GLP1.summary.means, GLP1.summary.se)
```

```{r barplot-Glucagon, dev=c('png','pdf')}
Glucagon.summary.means <- as.matrix(dcast(Glucagon.summary, Treatment~FeedingState, value.var="mean")[2:3])

Glucagon.summary.se <- as.matrix(dcast(Glucagon.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Glucagon.summary$mean + Glucagon.summary$se, na.rm=T)

plot <- barplot(Glucagon.summary.means,
                ylab="Serum Glucagon Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
legend("topleft", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, Glucagon.summary.means, Glucagon.summary.se)
```

```{r barplot-Ghrelin, dev=c('png','pdf')}
Ghrelin.summary.means <- as.matrix(dcast(Ghrelin.summary, Treatment~FeedingState, value.var="mean")[2:3])

Ghrelin.summary.se <- as.matrix(dcast(Ghrelin.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Ghrelin.summary$mean + Ghrelin.summary$se, na.rm=T)

plot <- barplot(Ghrelin.summary.means,
                ylab="Serum Ghrelin Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
legend("topright", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, Ghrelin.summary.means, Ghrelin.summary.se)
```

```{r barplot-Leptin, dev=c('png','pdf')}
Leptin.summary.means <- as.matrix(dcast(Leptin.summary, Treatment~FeedingState, value.var="mean")[2:3])

Leptin.summary.se <- as.matrix(dcast(Leptin.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Leptin.summary$mean + Leptin.summary$se, na.rm=T)

plot <- barplot(Leptin.summary.means,
                ylab="Serum Leptin Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
legend("topleft", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, Leptin.summary.means, Leptin.summary.se)
```

```{r barplot-Insulin, dev=c('png','pdf')}
Insulin.summary.means <- as.matrix(dcast(Insulin.summary, Treatment~FeedingState, value.var="mean")[2:3])

Insulin.summary.se <- as.matrix(dcast(Insulin.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Insulin.summary$mean + Insulin.summary$se, na.rm=T)

plot <- barplot(Insulin.summary.means,
                ylab="Serum Insulin Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
legend("topleft", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, Insulin.summary.means, Insulin.summary.se)
```

```{r barplot-Insulin.pmol.L, dev=c('png','pdf')}
Insulin.pmol.L.summary.means <- as.matrix(dcast(Insulin.pmol.L.summary, Treatment~FeedingState, value.var="mean")[2:3])

Insulin.pmol.L.summary.se <- as.matrix(dcast(Insulin.pmol.L.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Insulin.pmol.L.summary$mean + Insulin.pmol.L.summary$se, na.rm=T)

plot <- barplot(Insulin.pmol.L.summary.means,
                ylab="Serum Insulin Levels (pmol/L)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
legend("topleft", levels(data$Treatment), fill=Treatment.colors, bty="n")
superpose.eb(plot, Insulin.pmol.L.summary.means, Insulin.pmol.L.summary.se)
```

```{r barplot-all, dev=c('png','pdf')}
par(mfrow=c(2,4))
Resistin.summary.means <- as.matrix(dcast(Resistin.summary, Treatment~FeedingState, value.var="mean")[2:3])

Resistin.summary.se <- as.matrix(dcast(Resistin.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Resistin.summary$mean + Resistin.summary$se, na.rm=T)
superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

plot <- barplot(Resistin.summary.means,
                ylab="Serum Resistin Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:4],
                ylim = c(0,ymax),
                beside=T)
superpose.eb(plot, Resistin.summary.means, Resistin.summary.se)

GIP.summary.means <- as.matrix(dcast(GIP.summary, Treatment~FeedingState, value.var="mean")[2:3])

GIP.summary.se <- as.matrix(dcast(GIP.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(GIP.summary$mean + GIP.summary$se, na.rm=T)

plot <- barplot(GIP.summary.means,
                ylab="Serum GIP Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
superpose.eb(plot, GIP.summary.means, GIP.summary.se)

PAI1.summary.means <- as.matrix(dcast(PAI1.summary, Treatment~FeedingState, value.var="mean")[2:3])

PAI1.summary.se <- as.matrix(dcast(PAI1.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(PAI1.summary$mean + PAI1.summary$se, na.rm=T)

plot <- barplot(PAI1.summary.means,
                ylab="Serum PAI1 Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
superpose.eb(plot, PAI1.summary.means, PAI1.summary.se)

GLP1.summary.means <- as.matrix(dcast(GLP1.summary, Treatment~FeedingState, value.var="mean")[2:3])

GLP1.summary.se <- as.matrix(dcast(GLP1.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(GLP1.summary$mean + GLP1.summary$se, na.rm=T)

plot <- barplot(GLP1.summary.means,
                ylab="Serum GLP1 Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
superpose.eb(plot, GLP1.summary.means, GLP1.summary.se)

Glucagon.summary.means <- as.matrix(dcast(Glucagon.summary, Treatment~FeedingState, value.var="mean")[2:3])

Glucagon.summary.se <- as.matrix(dcast(Glucagon.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Glucagon.summary$mean + Glucagon.summary$se, na.rm=T)

plot <- barplot(Glucagon.summary.means,
                ylab="Serum Glucagon Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
superpose.eb(plot, Glucagon.summary.means, Glucagon.summary.se)

Ghrelin.summary.means <- as.matrix(dcast(Ghrelin.summary, Treatment~FeedingState, value.var="mean")[2:3])

Ghrelin.summary.se <- as.matrix(dcast(Ghrelin.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Ghrelin.summary$mean + Ghrelin.summary$se, na.rm=T)

plot <- barplot(Ghrelin.summary.means,
                ylab="Serum Ghrelin Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
superpose.eb(plot, Ghrelin.summary.means, Ghrelin.summary.se)

Leptin.summary.means <- as.matrix(dcast(Leptin.summary, Treatment~FeedingState, value.var="mean")[2:3])

Leptin.summary.se <- as.matrix(dcast(Leptin.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Leptin.summary$mean + Leptin.summary$se, na.rm=T)

plot <- barplot(Leptin.summary.means,
                ylab="Serum Leptin Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
superpose.eb(plot, Leptin.summary.means, Leptin.summary.se)

Insulin.summary.means <- as.matrix(dcast(Insulin.summary, Treatment~FeedingState, value.var="mean")[2:3])

Insulin.summary.se <- as.matrix(dcast(Insulin.summary, Treatment~FeedingState, value.var="se")[2:3])

ymax <- max(Insulin.summary$mean + Insulin.summary$se, na.rm=T)

plot <- barplot(Insulin.summary.means,
                ylab="Serum Insulin Levels (pg/mL)", 
                las=1,
                col = Treatment.colors[1:2],
                ylim = c(0,ymax),
                beside=T)
superpose.eb(plot, Insulin.summary.means, Insulin.summary.se)
```


# Session Information
```{r sessionInfo}
sessionInfo()
```