---
title: "Identifying and cleaning the comorbidity data for pneumonia"
author: "Dave Bridges"
date: "February 1, 2024"
output:
  html_document:
    highlight: tango
    keep_md: yes
    number_sections: yes
    toc: yes
---

## Purpose

To analyse the subset of patients with viral or bacterial pneumonia.  This script is to generate data about comorbidities.  This script can be found in `r getwd()` and was most recently run on `r date()`.

```{r input-data}
library(knitr)
#figures made will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures-cm/',
               echo=TRUE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)

library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(lubridate)

elixhauser.datafile <- 'ComorbiditiesElixhauserComprehensive.csv'
charlson.datafile <- 'ComorbiditiesCharlsonComprehensive.csv'
encounter.datafile <- 'EncounterAll.csv'
```

Combining demographic files

```{r combining}
#need to know when the first diagnosis was so need encounter data
encounter.data <- read_csv(encounter.datafile) %>%
  mutate(EncounterDate = as.Date(mdy_hm(DeID_AdmitDate),format="%y-%m-%d"))

cm.elix.data <- read_csv(elixhauser.datafile,na='-99') %>%
  left_join(encounter.data,by=c('DeID_PatientID','DeID_EncounterID')) %>%
  filter(!is.na(EncounterDate)) %>% #remove enounters without dates
  arrange(EncounterDate) %>%
  ungroup() %>%
  group_by(DeID_PatientID) %>%
  summarize(CardiacArrhythmias = max(CardiacArrhythmias,na.rm=T),
            CardiacArrhythmiasFirst = first(EncounterDate[CardiacArrhythmias==1]),
            HypertensionComplicated = max(HypertensionComplicated,na.rm=T),
            HypertensionComplicatedFirst = first(EncounterDate[HypertensionComplicated==1]),
            HypertensionUncomplicated = max(HypertensionUncomplicated,na.rm=T),
            HypertensionUncomplicatedFirst = first(EncounterDate[HypertensionUncomplicated==1]),
            LiverDisease = max(LiverDisease,na.rm=T),
            LiverDiseaseFirst = first(EncounterDate[LiverDisease==1]),
            Obesity = max(Obesity,na.rm=T),
            ObesityFirst = first(EncounterDate[Obesity==1]),
            RenalFailure = max(RenalFailure,na.rm=T),
            RenalFailureFirst = first(EncounterDate[RenalFailure==1]))


cm.charlson.data <- read_csv(charlson.datafile,na="-99") %>%
  left_join(encounter.data,by=c('DeID_PatientID','DeID_EncounterID')) %>%
  filter(!is.na(EncounterDate)) %>% #remove enounters without dates
  arrange(EncounterDate) %>%
  ungroup() %>%
  group_by(DeID_PatientID) %>%
  summarize(CerebrovascularDisease = max(CerebrovascularDisease,na.rm=T),
            CerebrovascularDiseaseFirst = first(EncounterDate[CerebrovascularDisease==1]),
            DiabetesWithChronicComplication = max(DiabetesWithChronicComplication,na.rm=T),
            DiabetesWithChronicComplicationFirst = first(EncounterDate[DiabetesWithChronicComplication==1]),
            DiabetesWithoutChronicComplication = max(DiabetesWithoutChronicComplication,na.rm=T),
            DiabetesWithoutChronicComplicationFirst = first(EncounterDate[DiabetesWithoutChronicComplication==1]))
```

# Separate Diagnoses

For COPD need to pull this information from the diagnosis summary file.

```{r diagnosis-copd}
diagnosis.datafile <- 'DiagnosesComprehensiveAll.csv'
diagnosis.data <- read_csv(diagnosis.datafile) 

copd.diagnoses <- 
  diagnosis.data %>%
  filter(grepl("OBSTRUCTIVE", TermNameSource)) %>%
  left_join(encounter.data, by=c("DeID_PatientID","DeID_EncounterID")) %>%
  arrange(EncounterDate) %>%
  distinct(DeID_PatientID,.keep_all = T) %>%
  mutate(COPD=1) %>%
  rename(COPDFirst=EncounterDate)
```

There were a total of `r nrow(copd.diagnoses)` patients with COPD diagnoses in the dataset.  These were combined with the other comorbidity data.

# Combining and Exporting

```{r combine-export}
combined.data <-
  full_join(cm.charlson.data,cm.elix.data,by="DeID_PatientID") %>%
  left_join(copd.diagnoses,by="DeID_PatientID")

comorbidities.filename <- 'Comorbidities_clean.csv'
write_csv(combined.data, comorbidities.filename)
```

Wrote this out to `r comorbidities.filename`.

# Session Information

```{r sessioninfo}
sessionInfo()
```