---
title: "Identifying the index cases for pneumonia"
author: "Dave Bridges"
date: "February 1, 2024"
output:
  html_document:
    highlight: tango
    keep_md: yes
    number_sections: yes
    toc: yes
---

## Purpose

To analyse the subset of patients with viral or bacterial pneumonia.  This script is to idenitify the index cases and ranges for when we are interested in outcomes.  This script can be found in `r getwd()` and was most recently run on `r date()`.

```{r input-data}
library(knitr)
#figures made will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures-index/',
               echo=TRUE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)

library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(lubridate)

diagnosis.datafile <- 'DiagnosesComprehensiveAll.csv'
encounter.datafile <- 'EncounterAll.csv'
encounter.anth.datafile <- "EncounterAnthropometricsBMI.csv"

```

```{r read-data}
library(readxl)
diagnosis.data <- read_csv(diagnosis.datafile) 
diagnosis.datasheet <- 'Pneumonia ICD9 and ICD10 Codes.xlsx'
mapped.diagnoses <- bind_rows(
  read_excel(diagnosis.datasheet, sheet="ICD9") %>% rename(ICD=1),
  read_excel(diagnosis.datasheet, sheet="ICD10", skip=1) %>% rename(ICD=1))%>% 
  select(ICD,Type)

bacterial.diagnoses <- filter(mapped.diagnoses, Type=="Bacterial") %>% pull(ICD)
viral.diagnoses <- filter(mapped.diagnoses, Type=="Viral") %>% pull(ICD)
other.diagnoses <- filter(mapped.diagnoses, Type=="Other") %>% pull(ICD)

pneumonia.diagnosis <- 
  diagnosis.data %>%
  filter(TermCodeSource %in% c(bacterial.diagnoses,viral.diagnoses,other.diagnoses)) %>% #there are also COPD diagnoses
  mutate(Type=case_when(TermCodeSource %in% bacterial.diagnoses ~ 'Bacterial',
                        TermCodeSource %in% viral.diagnoses ~ 'Viral',
                        TermCodeSource %in% other.diagnoses ~ 'Other',
                        .default=NA)) %>%
  filter(!is.na(Type)) #remove dignoses not in one of these categories

pneumonia.diagnosis %>%
  group_by(Type) %>%
  distinct(DeID_PatientID,.keep_all = T) %>%
  count %>%
  kable(caption="Diagnosis type by patient")
```

# Merging with Encounter Data

```{r encounter-merge}
encounter.data <- read_csv(encounter.datafile)
encounter.anthro.data <-read_csv(encounter.anth.datafile)

encounter.data.ano <- left_join(encounter.data,encounter.anthro.data,by=c("DeID_PatientID","DeID_EncounterID"))

#when was the encoutner for each pneumonia diagnosis
pneumonia.enc.data <-
  left_join(pneumonia.diagnosis,encounter.data.ano,by=c("DeID_PatientID","DeID_EncounterID")) %>%
  mutate(Duration=difftime(mdy_hm(DeID_DischargeDate),mdy_hm(DeID_AdmitDate),units="days")) #calculate duration between admit and discharge in days

patient.enc.data <- 
  pneumonia.enc.data %>%
  #select(DeID_PatientID,DeID_EncounterID,DeID_AdmitDate,DeID_DischargeDate,AgeInYears,EmergencyVisit,Duration,Type,BMI) %>%
  mutate(AdmitMonth=format_ISO8601(mdy_hm(DeID_AdmitDate),precision="ym")) %>% #calculated month and year of admit
  group_by(DeID_PatientID,AdmitMonth,Type) %>% #summarize encounters by patient,type and admit month, this is defined as one event, if its a different type of pneumonia its considered a new event
  summarize(AgeInYears=median(AgeInYears,na.rm=T), #year at the event
            BMI = median(BMI,na.rm=T), #bmi at the event
            EmergencyVisit=max(EmergencyVisit), # whether there was an emergency visit
            MaxStay=max(Duration), #the max of each duration within this event
            AdmitDate=min(format_ISO8601(mdy_hm(DeID_AdmitDate),precision="ymd")), #when they were admitted
            DischargeDate=max(format_ISO8601(mdy_hm(DeID_DischargeDate),precision="ymd"))) %>% #when they were discharged
  distinct(DeID_PatientID,AdmitMonth,Type,.keep_all = T) 

index.filename <- 'PneumoniaEncounterData.csv'
write_csv(patient.enc.data,index.filename)
```

After grouping multiple encounters into events there are `r nrow(patient.enc.data)` encounters from `r nrow(distinct(patient.enc.data,DeID_PatientID))`

This index encounter data is written out to `r index.filename` which contains data about the indexed encounter.

## Summary of Patient Encounters

```{r patient-encounters}
patient.enc.data %>% 
  ungroup %>%
  group_by(DeID_PatientID) %>%
  count(name="Admissions") %>%
  arrange(desc(Admissions)) ->
  patient.encounter.counts

kable(patient.encounter.counts %>% ungroup %>% count(Admissions,name="Number of Patients"))

library(ggplot2)
ggplot(patient.encounter.counts,(aes(x=Admissions))) +
  geom_histogram(stat='count',binwidth = 1) +
  labs(x="Admissions for Pneumonia",
       y="Number of Participants")

patient.enc.data %>% 
  ungroup %>%
  group_by(DeID_PatientID,Type) %>%
  count(name="Admissions") %>%
  arrange(desc(Admissions)) ->
  patient.encounter.type.counts

ggplot(patient.encounter.type.counts,(aes(x=Admissions))) +
  geom_histogram(stat='count',binwidth = 1) +
  facet_grid(~Type)+
  labs(x="Admissions for Pneumonia",
       y="Number of Participants")
```




# Session Information

```{r sessioninfo}
sessionInfo()
```