---
title: "Comparing PheCodes to ICD Codes"
author: "Dave Bridges"
date: "November 7, 2024"
format: html
editor: source
---

There was a differnce in case counts with our classification of pneumonia cases and the data on PheWeb.

After asking the data office learned this information

>We used the PheWAS package in R [PheWAS/PheWAS: The PheWAS R package](https://github.com/PheWAS/PheWAS) to create the phenotype cases and controls. This package mappings ICD9 and ICD10 codes to broader phenotypes called "phecodes." We used a minimum code count of 1 (i.e. an individual only had to have one of the ICD codes to be considered a case for the phenotype.) You can see which ICD codes map to each phenotype using the [Phecode Map v1.2](https://phewascatalog.org/phewas/#phe12). The two phenotypes you are interested in correspond to phecodes 480.1 and 480.11. The PheWAS package also applies exclusion ICD codes when determining controls – both phenotypes (480.1 and 480.11) exclude individuals with phecodes 480-488.99 from the control group. Additional details on phenotyping can be found in section 4 of our pheweb white paper: [PH_MGI_Freeze6_PheWeb](https://pheweb.org/MGI/static/PH_MGI_Freeze6_PheWeb_v1.1.pdf) 

## Purpose

To compare phecode-ICD9/10 mapping between our codes and the Phecodes and figure out if we have all the pertinent cases

## Data Entry

### ICD Code Data

These were obtained by inspecting the ICD codes and mapping them as bacterial or viral pneumonia.

```{r icd-codes, message=F}
library(readr)
library(readxl)
library(dplyr)
diagnosis.datasheet<- 'Pneumonia ICD9 and ICD10 Codes.xlsx'
icd.data <- mapped.diagnoses <- bind_rows(
  read_excel(diagnosis.datasheet, sheet="ICD9") %>% rename(ICD=1),
  read_excel(diagnosis.datasheet, sheet="ICD10", skip=1) %>% rename(ICD=1))%>% 
  select(ICD,Type)

bacterial.diagnoses <- filter(mapped.diagnoses, Type=="Bacterial") %>% pull(ICD)
viral.diagnoses <- filter(mapped.diagnoses, Type=="Viral") %>% pull(ICD)
other.diagnoses <- filter(mapped.diagnoses, Type=="Other") %>% pull(ICD)
```

We had used `r length(bacterial.diagnoses)` ICD9 and ICD10 codes in our analyses.

### Phecode Data

```{r phecode-mapping}
phecode.datafile <- 'https://phewascatalog.org/phewas/_w_bb5935e4/data/Phecode_map_v1_2_icd9_icd10cm.csv.zip'
phecode.destfile <- 'Phecode_map_v1_2_icd9_icd10cm_09_30_2024.csv.zip'
download.file(phecode.datafile,destfile=phecode.destfile)

phecode.data <- read_csv(phecode.destfile)
```

The phecode file was downlaoded from `r phecode.datafile`.

```{r phecode-pneumonia}
library(knitr)
filter(phecode.data, Phecode %in% c('480.1','480.11','480.12','480.13')) -> bacterial.pneumonia.phecodes 

bacterial.pneumonia.phecodes %>% kable(caption="ICD9/10 codes linked to Phecodes 480.1 and 480.11/2/3")

write(bacterial.pneumonia.phecodes$ICD, file = "Bacterial Pneumonia Phecodes.txt")

phecode.data %>% filter(PhecodeString=="Bacterial pneumonia") %>% kable(caption="ICD9/10 codes linked to bacterial pneumonia")
```

## Comparison of Codes

```{r comparason-codes}
# ICD not Phecode

filter(bacterial.pneumonia.phecodes, ICD %in% bacterial.diagnoses) %>% kable(caption="Phecodes we are using")

filter(bacterial.pneumonia.phecodes, !(ICD %in% bacterial.diagnoses)) %>% kable(caption="Phecodes derived ICD codes we are not using")

filter(mapped.diagnoses, Type=="Bacterial") %>%
  filter(!(ICD %in% c(bacterial.pneumonia.phecodes$ICD))) %>%
  left_join(phecode.data,by="ICD") -> icd.not.phecode

icd.not.phecode %>% kable(caption="ICD Codes we are using that are not in Phecodes")
```

### Summary

* There are `r nrow(filter(bacterial.pneumonia.phecodes, ICD %in% bacterial.diagnoses))` phecode-derived ICD codes that we are using
* There are `r nrow(filter(bacterial.pneumonia.phecodes, !(ICD %in% bacterial.diagnoses)))` phecode-derived ICD codes that we are not using
* There are `r nrow(icd.not.phecode)` ICD codes we are using that are not mapped as a phecode

```{r phecode-icd-venn}
library(ggVennDiagram)

# List of items
x <- list(Phecode = bacterial.pneumonia.phecodes$ICD, ICD=bacterial.diagnoses)

# 2D Venn diagram
library(ggplot2)
ggVennDiagram(x) 
```
