---
title: "Demographic Summary for Bacterial Pneumonia"
author: "Dave Bridges"
date: "February 1, 2024"
output:
  html_document:
    highlight: tango
    keep_md: yes
    number_sections: yes
    toc: yes
---

## Purpose

To analyse the subset of patients with viral or bacterial pneumonia.   This script combines the cleaned datasets and writes out a complete datafile for analyses.  This script can be found in `r getwd()` and was most recently run on `r date()`.

```{r input-data}
library(knitr)
#figures made will go to directory called figures, will make them as both png and pdf files 
opts_chunk$set(fig.path='figures-bacterial/',
               echo=TRUE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 2, digits = 3)

library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(lubridate)

complete.filename <- 'DatasetComplete.csv'
combined.data <- read_csv(complete.filename) %>%
  filter(Type=="Bacterial") %>%
  arrange(AdmitDate) %>%
  distinct(DeID_PatientID,.keep_all = T)
```



# Summary Demographics

```{r summary-dempographics}
quant.demo <- 
  combined.data %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(mean = ~mean(.,na.rm=T),
                             sd = ~sd(.,na.rm=T),
                             n = ~length(!(is.na(.)))))) %>%
  pivot_longer(cols=everything(),
               names_sep="_",names_to=c("Variable","Statistic")) %>%
  pivot_wider(names_from=Statistic,values_from=value)

kable(quant.demo, caption="Summary of quantitative values")

ances.demo.ql <- 
  combined.data %>%
  group_by(MajorityAncestry) %>%
  mutate(Type="Ancestry") %>%
  rename("Group"="MajorityAncestry") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ancestry")

type.demo.ql <- 
  combined.data %>%
  group_by(Type) %>%
  rename("Group"="Type") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Pneumonia Type")

emerg.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.x) %>%
  rename("Group"="EmergencyVisit.x") %>%
  count %>%
  mutate(Group=as.factor(Group)) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Emergency Visit") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

# coded as dates not yes/no
# cvd.demo.ql <- 
#   combined.data %>%
#   group_by(CerebrovascularDiseaseFirst) %>%
#   rename("Group"="CerebrovascularDiseaseFirst") %>%
#   count %>%
#   ungroup %>% 
#   mutate(Group=as.factor(Group)) %>%
#   mutate(Pct=n/sum(n)*100) %>%
#   mutate(Type="Cerebrovascular Disease")

gender.demo.ql <- 
  combined.data %>%
  group_by(GenderCode) %>%
  rename("Group"="GenderCode") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Gender")

race.demo.ql <- 
combined.data %>%
  group_by(RaceName) %>%
  count %>%
  ungroup %>%
  rename("Group"="RaceName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Race")

ethnicity.demo.ql <- 
combined.data %>%
  group_by(EthnicityName) %>%
  count %>%
  ungroup %>%
  rename("Group"="EthnicityName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ethnicity")

diabetes.demo.ql <- 
  combined.data %>%
  group_by(PriorDiabetes) %>%
  count %>%
  ungroup %>%
  rename("Group"="PriorDiabetes") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorDiabetes") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

smoking.demo.ql <- 
  combined.data %>%
  group_by(SmokingStatusMapped) %>%
  #mutate(Type="Smoking") %>%
  rename("Group"="SmokingStatusMapped") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Smoking") 

#update to hypertension first
hypertension.demo.ql <- 
  combined.data %>%
  group_by(PriorHypertension) %>%
  rename("Group"="PriorHypertension") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorHypertension") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

obesity.demo.ql <- 
  combined.data %>%
  group_by(PriorObesity) %>%
  rename("Group"="PriorObesity") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Obesity") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

copd.demo.ql <- 
  combined.data %>%
  group_by(PriorCOPD) %>%
  rename("Group"="PriorCOPD") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior COPD") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

arrythmia.demo.ql <- 
  combined.data %>%
  group_by(PriorCardiacArrhythmias) %>%
  rename("Group"="PriorCardiacArrhythmias") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Cardiac Arrhythmias") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

arrythmia.demo.ql <- 
  combined.data %>%
  group_by(PriorCardiacArrhythmias) %>%
  rename("Group"="PriorCardiacArrhythmias") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Cardiac Arrhythmias") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

death.demo.ql <- 
  combined.data %>%
  group_by(is.na(DeID_DeathIndexDeceasedDate)) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Death")  %>%
  mutate(Group=case_when(Group==TRUE~"Deceased",
                         Group==FALSE~"Alive"))

surv30d.demo.ql <- 
  combined.data %>%
  group_by(Survival.30day) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="30d Survival") 

surv60d.demo.ql <- 
  combined.data %>%
  group_by(Survival.60day) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="60d Survival") 

summary.discrete <-
  bind_rows(gender.demo.ql,
            race.demo.ql,
            ethnicity.demo.ql,
            ances.demo.ql,
            smoking.demo.ql,
            type.demo.ql,
            emerg.demo.ql, 
            death.demo.ql,
            surv30d.demo.ql,
            surv60d.demo.ql,
            diabetes.demo.ql,
            hypertension.demo.ql,
            arrythmia.demo.ql,
            obesity.demo.ql,
            copd.demo.ql
            ) %>%
  select(Type,Group,n,Pct)

kable(summary.discrete, caption="Summary of discrete variables for all pneumonia cases")
```

# By 30 Day Survival

```{r 30d-survival}
combined.data <- 
  combined.data %>%
  mutate(Survival.30day.Group = as.factor(case_when(Survival.30day=="Within 30 Days" ~ "No",
                                    Survival.30day!="Survival Past 30 Days" ~ "Yes",
                                    is.na(Survival.30day) ~ "Yes",
                                    .default="Yes")))

quant.demo <- 
  combined.data %>%
  group_by(Survival.30day.Group) %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(mean = ~mean(.,na.rm=T),
                             sd = ~sd(.,na.rm=T),
                             #shapiro.p = ~shapiro.test(.)$p.value,
                             n = ~length(!(is.na(.)))))) %>%
  pivot_longer(cols=contains("_"), #fix this
               names_sep="_",names_to=c("Variable","Statistic")) %>%
  pivot_wider(names_from=Statistic,values_from=value) %>%
  arrange(Variable)

kable(quant.demo, caption="Summary of quantitative values")
# wilcoxon tests, not normally distributed

quant.t.tests <-
  combined.data %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(Mann.Whitney=~wilcox.test(.~Survival.30day.Group)$p.value)))

kable(quant.t.tests,caption="Mann-Whitney tests for 30 day survival",digits=c(99,99,99))

ances.demo.ql <- 
  combined.data %>%
  group_by(MajorityAncestry,Survival.30day.Group) %>%
  mutate(Type="Ancestry") %>%
  rename("Group"="MajorityAncestry") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ancestry")

type.demo.ql <- 
  combined.data %>%
  group_by(Type,Survival.30day.Group) %>%
  rename("Group"="Type") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Pneumonia Type")

emerg.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.x,Survival.30day.Group) %>%
  rename("Group"="EmergencyVisit.x") %>%
  count %>%
  mutate(Group=as.factor(Group)) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Emergency Visit") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

# coded as dates not yes/no
# cvd.demo.ql <- 
#   combined.data %>%
#   group_by(CerebrovascularDiseaseFirst) %>%
#   rename("Group"="CerebrovascularDiseaseFirst") %>%
#   count %>%
#   ungroup %>% 
#   mutate(Group=as.factor(Group)) %>%
#   mutate(Pct=n/sum(n)*100) %>%
#   mutate(Type="Cerebrovascular Disease")

gender.demo.ql <- 
  combined.data %>%
  group_by(GenderCode,Survival.30day.Group) %>%
  rename("Group"="GenderCode") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Gender")

race.demo.ql <- 
combined.data %>%
  group_by(RaceName,Survival.30day.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="RaceName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Race")

ethnicity.demo.ql <- 
combined.data %>%
  group_by(EthnicityName,Survival.30day.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="EthnicityName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ethnicity")

diabetes.demo.ql <- 
  combined.data %>%
  group_by(PriorDiabetes,Survival.30day.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="PriorDiabetes") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorDiabetes") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

smoking.demo.ql <- 
  combined.data %>%
  group_by(SmokingStatusMapped,Survival.30day.Group) %>%
  #mutate(Type="Smoking") %>%
  rename("Group"="SmokingStatusMapped") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Smoking") 

#update to hypertension first
hypertension.demo.ql <- 
  combined.data %>%
  group_by(PriorHypertension,Survival.30day.Group) %>%
  rename("Group"="PriorHypertension") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorHypertension") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

arrythmia.demo.ql <- 
  combined.data %>%
  group_by(PriorCardiacArrhythmias,Survival.30day.Group) %>%
  rename("Group"="PriorCardiacArrhythmias") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Cardiac Arrhythmias") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

obesity.demo.ql <- 
  combined.data %>%
  group_by(PriorObesity,Survival.30day.Group) %>%
  rename("Group"="PriorObesity") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Obesity") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

copd.demo.ql <- 
  combined.data %>%
  group_by(PriorCOPD,Survival.30day.Group) %>%
  rename("Group"="PriorCOPD") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior COPD") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

death.demo.ql <- 
  combined.data %>%
  group_by(is.na(DeID_DeathIndexDeceasedDate),Survival.30day.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Death")  %>%
  mutate(Group=case_when(Group==TRUE~"Deceased",
                         Group==FALSE~"Alive"))

surv30d.demo.ql <- 
  combined.data %>%
  group_by(Survival.30day,Survival.30day.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="30d Survival") 

surv60d.demo.ql <- 
  combined.data %>%
  group_by(Survival.60day,Survival.30day.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="60d Survival") 

total.survival <- combined.data %>% group_by(Survival.30day.Group) %>% count() %>% pull(n)

summary.discrete <-
  bind_rows(gender.demo.ql,
            race.demo.ql,
            ethnicity.demo.ql,
            ances.demo.ql,
            smoking.demo.ql,
            type.demo.ql,
            emerg.demo.ql, 
            death.demo.ql,
            surv30d.demo.ql,
            surv60d.demo.ql,
            diabetes.demo.ql,
            hypertension.demo.ql,
            arrythmia.demo.ql,
            obesity.demo.ql,
            copd.demo.ql
            ) %>%
  select(Survival.30day.Group,Type,Group,n,Pct) %>%
  ungroup %>%
  select(-Pct) %>%
  pivot_wider(names_from = Survival.30day.Group,values_from = n, values_fn = sum) %>%
  arrange(Type,Group) %>%
  replace(.=="NULL", NA) %>%
  mutate(No = case_when(is.na(No)~0,
                            !(is.na(No))~No)) %>%
  mutate(Yes = case_when(is.na(Yes)~0,
                            !(is.na(Yes))~Yes)) %>%
  rowwise %>%
  mutate(Pct=No/(No+Yes)*100) %>%
  mutate(Chisq.p = chisq.test(x=c(No,Yes),p=total.survival,rescale.p = T)$p.value) %>%
  mutate(Sig = case_when(Chisq.p < 0.01 ~ '**',
                         Chisq.p < 0.05 ~ '*'))

kable(summary.discrete, 
      caption=paste("Summary of discrete variables for bacterial pneumonia cases, stratified by 30 day survival.  Overall", round(total.survival[1]/(total.survival[2]+total.survival[1])*100,2),"% visited died at 30d"),
      digits=c(0,0,0,0,1,99,0))
```

# By 60 Day Survival

```{r 60d-survival}
combined.data <- 
  combined.data %>%
  mutate(Survival.60.day.Group = case_when(Survival.60day=="Within 60 Days" ~ "No",
                                    Survival.60day!="Survival Past 60 Days" ~ "Yes",
                                    is.na(Survival.60day) ~ "Yes",
                                    .default="Yes"))

quant.demo <- 
  combined.data %>%
  group_by(Survival.60.day.Group) %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(mean = ~mean(.,na.rm=T),
                             sd = ~sd(.,na.rm=T),
                             #shapiro.p = ~shapiro.test(.)$p.value,
                             n = ~length(!(is.na(.)))))) %>%
  pivot_longer(cols=contains("_"), #all variables separated by underscore
               names_sep="_",names_to=c("Variable","Statistic")) %>%
  pivot_wider(names_from=Statistic,values_from=value) %>%
  arrange(Variable)

kable(quant.demo, caption="Summary of quantitative values")

quant.t.tests <-
  combined.data %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(Mann.Whitney=~wilcox.test(.~Survival.30day.Group)$p.value)))

kable(quant.t.tests,captionn="Mann-Whitney tests for 30 day survival", digits=c(99,99,99))

ances.demo.ql <- 
  combined.data %>%
  group_by(MajorityAncestry,Survival.60.day.Group) %>%
  mutate(Type="Ancestry") %>%
  rename("Group"="MajorityAncestry") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ancestry")

type.demo.ql <- 
  combined.data %>%
  group_by(Type,Survival.60.day.Group) %>%
  rename("Group"="Type") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Pneumonia Type")

emerg.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.x,Survival.60.day.Group) %>%
  rename("Group"="EmergencyVisit.x") %>%
  count %>%
  mutate(Group=as.factor(Group)) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Emergency Visit") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

# coded as dates not yes/no
# cvd.demo.ql <- 
#   combined.data %>%
#   group_by(CerebrovascularDiseaseFirst) %>%
#   rename("Group"="CerebrovascularDiseaseFirst") %>%
#   count %>%
#   ungroup %>% 
#   mutate(Group=as.factor(Group)) %>%
#   mutate(Pct=n/sum(n)*100) %>%
#   mutate(Type="Cerebrovascular Disease")

gender.demo.ql <- 
  combined.data %>%
  group_by(GenderCode,Survival.60.day.Group) %>%
  rename("Group"="GenderCode") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Gender")

race.demo.ql <- 
combined.data %>%
  group_by(RaceName,Survival.60.day.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="RaceName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Race")

ethnicity.demo.ql <- 
combined.data %>%
  group_by(EthnicityName,Survival.60.day.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="EthnicityName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ethnicity")

diabetes.demo.ql <- 
  combined.data %>%
  group_by(PriorDiabetes,Survival.60.day.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="PriorDiabetes") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorDiabetes") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

smoking.demo.ql <- 
  combined.data %>%
  group_by(SmokingStatusMapped,Survival.60.day.Group) %>%
  #mutate(Type="Smoking") %>%
  rename("Group"="SmokingStatusMapped") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Smoking") 

#update to hypertension first
hypertension.demo.ql <- 
  combined.data %>%
  group_by(PriorHypertension,Survival.60.day.Group) %>%
  rename("Group"="PriorHypertension") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorHypertension") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

arrythmia.demo.ql <- 
  combined.data %>%
  group_by(PriorCardiacArrhythmias,Survival.60.day.Group) %>%
  rename("Group"="PriorCardiacArrhythmias") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Cardiac Arrhythmias") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

obesity.demo.ql <- 
  combined.data %>%
  group_by(PriorObesity,Survival.60.day.Group) %>%
  rename("Group"="PriorObesity") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Obesity") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

copd.demo.ql <- 
  combined.data %>%
  group_by(PriorCOPD,Survival.60.day.Group) %>%
  rename("Group"="PriorCOPD") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior COPD") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

death.demo.ql <- 
  combined.data %>%
  group_by(is.na(DeID_DeathIndexDeceasedDate),Survival.60.day.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Death")  %>%
  mutate(Group=case_when(Group==TRUE~"Deceased",
                         Group==FALSE~"Alive"))

surv30d.demo.ql <- 
  combined.data %>%
  group_by(Survival.30day,Survival.60.day.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="60d Survival") 

surv60d.demo.ql <- 
  combined.data %>%
  group_by(Survival.60day,Survival.60.day.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="60d Survival") 

total.survival <- combined.data %>% group_by(Survival.60.day.Group) %>% count() %>% pull(n)

summary.discrete <-
  bind_rows(gender.demo.ql,
            race.demo.ql,
            ethnicity.demo.ql,
            ances.demo.ql,
            smoking.demo.ql,
            type.demo.ql,
            emerg.demo.ql, 
            death.demo.ql,
            surv30d.demo.ql,
            surv60d.demo.ql,
            diabetes.demo.ql,
            hypertension.demo.ql,
            arrythmia.demo.ql,
            obesity.demo.ql,
            copd.demo.ql
            ) %>%
  select(Survival.60.day.Group,Type,Group,n,Pct) %>%
  ungroup %>%
  select(-Pct) %>%
  pivot_wider(names_from = Survival.60.day.Group,values_from = n, values_fn = sum) %>%
  arrange(Type,Group) %>%
  replace(.=="NULL", NA) %>%
  mutate(No = case_when(is.na(No)~0,
                            !(is.na(No))~No)) %>%
  mutate(Yes = case_when(is.na(Yes)~0,
                            !(is.na(Yes))~Yes)) %>%
  rowwise %>%
  mutate(Pct=No/(No+Yes)*100) %>% #because want not survived
  mutate(Chisq.p = chisq.test(x=c(No,Yes),p=total.survival,rescale.p = T)$p.value) %>%
  mutate(Sig = case_when(Chisq.p < 0.01 ~ '**',
                         Chisq.p < 0.05 ~ '*'))

kable(summary.discrete, 
      caption=paste("Summary of discrete variables for bacterial pneumonia cases, stratified by 60 day survival.  Overall", round(total.survival[1]/(total.survival[2]+total.survival[1])*100,2),"% visited died at 60d"),
      digits=c(0,0,0,0,1,99,0))
```

# By ER Visit

```{r er-visit}
combined.data <- 
  combined.data %>%
  mutate(EmergencyVisit.Group = case_when(EmergencyVisit.x==0 ~ "No",
                                    EmergencyVisit.x==1 ~ "Yes",
                                    is.na(EmergencyVisit.x) ~ "No",
                                    .default="No"))

quant.demo <- 
  combined.data %>%
  group_by(EmergencyVisit.Group) %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(mean = ~mean(.,na.rm=T),
                             sd = ~sd(.,na.rm=T),
                             #shapiro.p = ~shapiro.test(.)$p.value,
                             n = ~length(!(is.na(.)))))) %>%
  pivot_longer(cols=contains("_"),#all varaibles separated by underscore
               names_sep="_",names_to=c("Variable","Statistic")) %>%
  pivot_wider(names_from=Statistic,values_from=value) %>%
  arrange(Variable)

kable(quant.demo, caption="Summary of quantitative values")
# wilcoxon tests, not normally distributed

quant.t.tests <-
  combined.data %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(Mann.Whitney=~wilcox.test(.~EmergencyVisit.Group)$p.value)))

kable(quant.t.tests,captionn="Mann-Whitney tests for ER visit",digits=c(99,99,99))

ances.demo.ql <- 
  combined.data %>%
  group_by(MajorityAncestry,EmergencyVisit.Group) %>%
  mutate(Type="Ancestry") %>%
  rename("Group"="MajorityAncestry") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ancestry")

type.demo.ql <- 
  combined.data %>%
  group_by(Type,EmergencyVisit.Group) %>%
  rename("Group"="Type") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Pneumonia Type")

emerg.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.Group) %>%
  rename("Group"="EmergencyVisit.Group") %>%
  count %>%
  mutate(Group=as.factor(Group)) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Emergency Visit") 

# coded as dates not yes/no
# cvd.demo.ql <- 
#   combined.data %>%
#   group_by(CerebrovascularDiseaseFirst) %>%
#   rename("Group"="CerebrovascularDiseaseFirst") %>%
#   count %>%
#   ungroup %>% 
#   mutate(Group=as.factor(Group)) %>%
#   mutate(Pct=n/sum(n)*100) %>%
#   mutate(Type="Cerebrovascular Disease")

gender.demo.ql <- 
  combined.data %>%
  group_by(GenderCode,EmergencyVisit.Group) %>%
  rename("Group"="GenderCode") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Gender")

race.demo.ql <- 
combined.data %>%
  group_by(RaceName,EmergencyVisit.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="RaceName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Race")

ethnicity.demo.ql <- 
combined.data %>%
  group_by(EthnicityName,EmergencyVisit.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="EthnicityName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ethnicity")

diabetes.demo.ql <- 
  combined.data %>%
  group_by(PriorDiabetes,EmergencyVisit.Group) %>%
  count %>%
  ungroup %>%
  rename("Group"="PriorDiabetes") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorDiabetes") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

smoking.demo.ql <- 
  combined.data %>%
  group_by(SmokingStatusMapped,EmergencyVisit.Group) %>%
  #mutate(Type="Smoking") %>%
  rename("Group"="SmokingStatusMapped") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Smoking") 

#update to hypertension first
hypertension.demo.ql <- 
  combined.data %>%
  group_by(PriorHypertension,EmergencyVisit.Group) %>%
  rename("Group"="PriorHypertension") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorHypertension") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

arrythmia.demo.ql <- 
  combined.data %>%
  group_by(PriorCardiacArrhythmias,EmergencyVisit.Group) %>%
  rename("Group"="PriorCardiacArrhythmias") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Cardiac Arrhythmias") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

obesity.demo.ql <- 
  combined.data %>%
  group_by(PriorObesity,EmergencyVisit.Group) %>%
  rename("Group"="PriorObesity") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Obesity") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

copd.demo.ql <- 
  combined.data %>%
  group_by(PriorCOPD,EmergencyVisit.Group) %>%
  rename("Group"="PriorCOPD") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior COPD") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

death.demo.ql <- 
  combined.data %>%
  group_by(is.na(DeID_DeathIndexDeceasedDate),EmergencyVisit.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Death")  %>%
  mutate(Group=case_when(Group==TRUE~"Deceased",
                         Group==FALSE~"Alive"))

surv30d.demo.ql <- 
  combined.data %>%
  group_by(Survival.30day,EmergencyVisit.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="30d Survival") 

surv60d.demo.ql <- 
  combined.data %>%
  group_by(Survival.60day,EmergencyVisit.Group) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="60d Survival") 

total.survival <- combined.data %>% group_by(EmergencyVisit.Group) %>% count() %>% pull(n)

summary.discrete <-
  bind_rows(gender.demo.ql,
            race.demo.ql,
            ethnicity.demo.ql,
            ances.demo.ql,
            smoking.demo.ql,
            type.demo.ql,
            #emerg.demo.ql, 
            death.demo.ql,
            surv30d.demo.ql,
            surv60d.demo.ql,
            diabetes.demo.ql,
            hypertension.demo.ql,
            arrythmia.demo.ql,
            obesity.demo.ql,
            copd.demo.ql
            ) %>%
  select(EmergencyVisit.Group,Type,Group,n,Pct) %>%
  ungroup %>%
  select(-Pct) %>%
  pivot_wider(names_from = EmergencyVisit.Group,values_from = n, values_fn = sum) %>%
  arrange(Type,Group) %>%
  replace(.=="NULL", NA) %>%
  mutate(No = case_when(is.na(No)~0,
                            !(is.na(No))~No)) %>%
  mutate(Yes = case_when(is.na(Yes)~0,
                            !(is.na(Yes))~Yes)) %>%
  rowwise %>%
  mutate(Pct=Yes/(No+Yes)*100) %>%
  mutate(Chisq.p = chisq.test(x=c(No,Yes),p=total.survival,rescale.p = T)$p.value) %>%
  mutate(Sig = case_when(Chisq.p < 0.01 ~ '**',
                         Chisq.p < 0.05 ~ '*'))

kable(summary.discrete, 
      caption=paste("Summary of discrete variables for bacterial pneumonia cases, stratified by ER visit survival.  Overall", round(total.survival[2]/(total.survival[2]+total.survival[1])*100,2),"% visited ER"),
      digits=c(0,0,0,0,1,99,0))
```

# By Longer than Exoected Stay

The median stay was `r median(combined.data$MaxStay,na.rm=T)`.  We set the a longer than average stay to be **3 days**.

```{r length-of-stay}
combined.data <- 
  combined.data %>%
  mutate(LongStay=case_when(MaxStay>=3~"Yes",
                                MaxStay<3~"No"))

quant.demo <- 
  combined.data %>%
  group_by(LongStay) %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(mean = ~mean(.,na.rm=T),
                             sd = ~sd(.,na.rm=T),
                             #shapiro.p = ~shapiro.test(.)$p.value,
                             n = ~length(!(is.na(.)))))) %>%
  pivot_longer(cols=2:10,
               names_sep="_",names_to=c("Variable","Statistic")) %>%
  pivot_wider(names_from=Statistic,values_from=value) %>%
  arrange(Variable)

kable(quant.demo, caption="Summary of quantitative values for above average length of stay")
# wilcoxon tests, not normally distributed

quant.t.tests <-
  combined.data %>%
  summarize(across(.cols=c('AgeInYears.x','BMI'),
                   .fns=list(Mann.Whitney=~wilcox.test(.~LongStay)$p.value)))

kable(quant.t.tests,captionn="Mann-Whitney tests for ER visit",digits=c(99,99,99))

ances.demo.ql <- 
  combined.data %>%
  group_by(MajorityAncestry,LongStay) %>%
  mutate(Type="Ancestry") %>%
  rename("Group"="MajorityAncestry") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ancestry")

type.demo.ql <- 
  combined.data %>%
  group_by(Type,LongStay) %>%
  rename("Group"="Type") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Pneumonia Type")

emerg.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.Group,LongStay) %>%
  rename("Group"="EmergencyVisit.Group") %>%
  count %>%
  mutate(Group=as.factor(Group)) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Emergency Visit") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

# coded as dates not yes/no
# cvd.demo.ql <- 
#   combined.data %>%
#   group_by(CerebrovascularDiseaseFirst) %>%
#   rename("Group"="CerebrovascularDiseaseFirst") %>%
#   count %>%
#   ungroup %>% 
#   mutate(Group=as.factor(Group)) %>%
#   mutate(Pct=n/sum(n)*100) %>%
#   mutate(Type="Cerebrovascular Disease")

gender.demo.ql <- 
  combined.data %>%
  group_by(GenderCode,LongStay) %>%
  rename("Group"="GenderCode") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Gender")

race.demo.ql <- 
combined.data %>%
  group_by(RaceName,LongStay) %>%
  count %>%
  ungroup %>%
  rename("Group"="RaceName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Race")

ethnicity.demo.ql <- 
combined.data %>%
  group_by(EthnicityName,LongStay) %>%
  count %>%
  ungroup %>%
  rename("Group"="EthnicityName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ethnicity")

diabetes.demo.ql <- 
  combined.data %>%
  group_by(PriorDiabetes,LongStay) %>%
  count %>%
  ungroup %>%
  rename("Group"="PriorDiabetes") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorDiabetes") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

smoking.demo.ql <- 
  combined.data %>%
  group_by(SmokingStatusMapped,LongStay) %>%
  #mutate(Type="Smoking") %>%
  rename("Group"="SmokingStatusMapped") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Smoking") 

#update to hypertension first
hypertension.demo.ql <- 
  combined.data %>%
  group_by(PriorHypertension,LongStay) %>%
  rename("Group"="PriorHypertension") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorHypertension") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

arrythmia.demo.ql <- 
  combined.data %>%
  group_by(PriorCardiacArrhythmias,LongStay) %>%
  rename("Group"="PriorCardiacArrhythmias") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Cardiac Arrhythmias") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

obesity.demo.ql <- 
  combined.data %>%
  group_by(PriorObesity,LongStay) %>%
  rename("Group"="PriorObesity") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Obesity") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

copd.demo.ql <- 
  combined.data %>%
  group_by(PriorCOPD,LongStay) %>%
  rename("Group"="PriorCOPD") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior COPD") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

death.demo.ql <- 
  combined.data %>%
  group_by(is.na(DeID_DeathIndexDeceasedDate),LongStay) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Death")  %>%
  mutate(Group=case_when(Group==TRUE~"Deceased",
                         Group==FALSE~"Alive"))

surv30d.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.Group,LongStay) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="30d Survival") 

surv60d.demo.ql <- 
  combined.data %>%
  group_by(Survival.60day,LongStay) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="60d Survival") 

total.survival <- combined.data %>% group_by(LongStay) %>% count() %>% pull(n) 

summary.discrete <-
  bind_rows(gender.demo.ql,
            race.demo.ql,
            ethnicity.demo.ql,
            ances.demo.ql,
            smoking.demo.ql,
            type.demo.ql,
            emerg.demo.ql, 
            death.demo.ql,
            surv30d.demo.ql %>% mutate(Group=as.character(Group)),
            surv60d.demo.ql,
            diabetes.demo.ql,
            hypertension.demo.ql,
            arrythmia.demo.ql,
            obesity.demo.ql,
            copd.demo.ql
            ) %>%
  select(LongStay,Type,Group,n,Pct) %>%
  ungroup %>%
  select(-Pct) %>%
  pivot_wider(names_from = LongStay,values_from = n, values_fn = sum) %>%
  arrange(Type,Group) %>%
  replace(.=="NULL", NA) %>%
  mutate(No = case_when(is.na(No)~0,
                            !(is.na(No))~No)) %>%
  mutate(Yes = case_when(is.na(Yes)~0,
                            !(is.na(Yes))~Yes)) %>%
  rowwise %>%
  mutate(Pct=Yes/(No+Yes)*100) %>%
  filter(No+Yes>0) %>%
  mutate(Chisq.p = chisq.test(x=c(No,Yes),p=total.survival[1:2],rescale.p = T)$p.value) %>%
  mutate(Sig = case_when(Chisq.p < 0.01 ~ '**',
                         Chisq.p < 0.05 ~ '*'))

kable(summary.discrete, 
      caption="Summary of discrete variables for bacterial pneumonia cases, stratified by above average length of stay",
      digits=c(0,0,0,0,1,99,0))
```



# By Need for Ventillation


```{r ventillation}
procedure.file <- 'ProceduresComprehensive.csv'
procedure.data <- read_csv(procedure.file)

combined.data <- 
  combined.data %>%
  mutate(Ventillation=if_else(DeID_PatientID %in% procedure.data$DeID_PatientID,
                              "Yes",
                              "No"))

quant.demo <- 
  combined.data %>%
  group_by(Ventillation) %>%
  summarize(across(.cols=c('AgeInYears.x','BMI','MaxStay'),
                   .fns=list(mean = ~mean(.,na.rm=T),
                             sd = ~sd(.,na.rm=T),
                             #shapiro.p = ~shapiro.test(.)$p.value,
                             n = ~length(!(is.na(.)))))) %>%
  pivot_longer(cols=2:10,
               names_sep="_",names_to=c("Variable","Statistic")) %>%
  pivot_wider(names_from=Statistic,values_from=value) %>%
  arrange(Variable)

kable(quant.demo, caption="Summary of quantitative values for the need for ventillation")
# wilcoxon tests, not normally distributed

quant.t.tests <-
  combined.data %>%
  summarize(across(.cols=c('AgeInYears.x','BMI'),
                   .fns=list(Mann.Whitney=~wilcox.test(.~Ventillation)$p.value)))

kable(quant.t.tests,captionn="Mann-Whitney tests for need for ventillation",digits=c(99,99,99))

ances.demo.ql <- 
  combined.data %>%
  group_by(MajorityAncestry,Ventillation) %>%
  mutate(Type="Ancestry") %>%
  rename("Group"="MajorityAncestry") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ancestry")

type.demo.ql <- 
  combined.data %>%
  group_by(Type,Ventillation) %>%
  rename("Group"="Type") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Pneumonia Type")

emerg.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.Group,Ventillation) %>%
  rename("Group"="EmergencyVisit.Group") %>%
  count %>%
  mutate(Group=as.factor(Group)) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Emergency Visit") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

# coded as dates not yes/no
# cvd.demo.ql <- 
#   combined.data %>%
#   group_by(CerebrovascularDiseaseFirst) %>%
#   rename("Group"="CerebrovascularDiseaseFirst") %>%
#   count %>%
#   ungroup %>% 
#   mutate(Group=as.factor(Group)) %>%
#   mutate(Pct=n/sum(n)*100) %>%
#   mutate(Type="Cerebrovascular Disease")

gender.demo.ql <- 
  combined.data %>%
  group_by(GenderCode,Ventillation) %>%
  rename("Group"="GenderCode") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Gender")

race.demo.ql <- 
combined.data %>%
  group_by(RaceName,Ventillation) %>%
  count %>%
  ungroup %>%
  rename("Group"="RaceName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Race")

ethnicity.demo.ql <- 
combined.data %>%
  group_by(EthnicityName,Ventillation) %>%
  count %>%
  ungroup %>%
  rename("Group"="EthnicityName") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Ethnicity")

diabetes.demo.ql <- 
  combined.data %>%
  group_by(PriorDiabetes,Ventillation) %>%
  count %>%
  ungroup %>%
  rename("Group"="PriorDiabetes") %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorDiabetes") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

smoking.demo.ql <- 
  combined.data %>%
  group_by(SmokingStatusMapped,Ventillation) %>%
  #mutate(Type="Smoking") %>%
  rename("Group"="SmokingStatusMapped") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Smoking") 

#update to hypertension first
hypertension.demo.ql <- 
  combined.data %>%
  group_by(PriorHypertension,Ventillation) %>%
  rename("Group"="PriorHypertension") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="PriorHypertension") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

arrythmia.demo.ql <- 
  combined.data %>%
  group_by(PriorCardiacArrhythmias,Ventillation) %>%
  rename("Group"="PriorCardiacArrhythmias") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Cardiac Arrhythmias") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

obesity.demo.ql <- 
  combined.data %>%
  group_by(PriorObesity,Ventillation) %>%
  rename("Group"="PriorObesity") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior Obesity") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

copd.demo.ql <- 
  combined.data %>%
  group_by(PriorCOPD,Ventillation) %>%
  rename("Group"="PriorCOPD") %>%
  count %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Prior COPD") %>%
  mutate(Group=case_when(Group==1~"Yes",
                         Group==2~"No"))

death.demo.ql <- 
  combined.data %>%
  group_by(is.na(DeID_DeathIndexDeceasedDate),Ventillation) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="Death")  %>%
  mutate(Group=case_when(Group==TRUE~"Deceased",
                         Group==FALSE~"Alive"))

surv30d.demo.ql <- 
  combined.data %>%
  group_by(EmergencyVisit.Group,Ventillation) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="30d Survival") 

surv60d.demo.ql <- 
  combined.data %>%
  group_by(Survival.60day,Ventillation) %>%
  count() %>%
  rename("Group"=1) %>%
  ungroup %>%
  mutate(Pct=n/sum(n)*100) %>%
  mutate(Type="60d Survival") 

total.survival <- combined.data %>% group_by(Ventillation) %>% count() %>% pull(n) 

summary.discrete <-
  bind_rows(gender.demo.ql,
            race.demo.ql,
            ethnicity.demo.ql,
            ances.demo.ql,
            smoking.demo.ql,
            type.demo.ql,
            emerg.demo.ql, 
            death.demo.ql,
            surv30d.demo.ql %>% mutate(Group=as.character(Group)),
            surv60d.demo.ql,
            diabetes.demo.ql,
            hypertension.demo.ql,
            arrythmia.demo.ql,
            obesity.demo.ql,
            copd.demo.ql
            ) %>%
  select(Ventillation,Type,Group,n,Pct) %>%
  ungroup %>%
  select(-Pct) %>%
  pivot_wider(names_from = Ventillation,values_from = n, values_fn = sum) %>%
  arrange(Type,Group) %>%
  replace(.=="NULL", NA) %>%
  mutate(No = case_when(is.na(No)~0,
                            !(is.na(No))~No)) %>%
  mutate(Yes = case_when(is.na(Yes)~0,
                            !(is.na(Yes))~Yes)) %>%
  rowwise %>%
  mutate(Pct=Yes/(No+Yes)*100) %>%
  filter(No+Yes>0) %>%
  mutate(Chisq.p = chisq.test(x=c(No,Yes),p=total.survival[1:2],rescale.p = T)$p.value) %>%
  mutate(Sig = case_when(Chisq.p < 0.01 ~ '**',
                         Chisq.p < 0.05 ~ '*'))

kable(summary.discrete, 
      caption=paste("Summary of discrete variables for bacterial pneumonia cases, stratified by ventillation status.  Overall", round(total.survival[2]/(total.survival[2]+total.survival[1])*100,2),"% required ventillation"),
      digits=c(0,0,0,0,1,99,0))
```

```{r output}
genetics_outfile <- '../Case_Control_After_Bacterial_Pneumonia.csv'

combined.data %>%
  rename('Age'='AgeInYears.x') %>%
  mutate(AgeOver65 = case_when(Age>=65~1,
                               Age<65~0)) %>%
  mutate(COPD.History = case_when(COPD==1~1,
                                  .default=0)) %>% 
  select(DeID_PatientID,
         Age,AgeOver65,GenderCode,BMI,
         RaceName,EthnicityName,
         SmokingStatusMapped,COPD.History,Diabetes,
         Survival.30day, Survival.60day, EmergencyVisit.Group, Ventillation,LongStay,
         SubType) %>% 
  write_csv(genetics_outfile)
```



# Session Information

```{r sessioninfo}
sessionInfo()
```